---
title: "Analysis - SMC Paper"
author: "Tom Offrede"
date: "2023-01-26"
output:
  html_document:
    toc: true
    toc_float: 
      collapsed: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(lme4)
library(tidyverse)

folder <- "C:/Users/offredet/Documents/1HU/ExperimentBreathing/Data/DataForAnalysis/AllData/"

orderCond <- c("Sitting", "Light", "Heavy")
orderCondBase <- c("Baseline", "Sitting", "Light", "Heavy")
orderAct <- c("watching", "listening")

load(paste0(folder, "DataBreathing.RData"))
load(paste0(folder, "DataReadSpeech.RData"))
load(paste0(folder, "DataSpeech.RData"))

brm <- brm %>% 
  mutate_at(c("Task", "act", "Condition", "Speaker"), as.factor) %>% 
  mutate(across(Condition, factor, levels=c("Baseline", "Sitting","Light","Heavy")),
         across(act, factor, levels=c("watching","listening", "speaking")),
         Cond2 = ifelse(Condition == "Baseline", "Baseline", "Interaction"))

frb <- frb %>% 
  mutate_at(c("Task", "Condition", "Speaker"), as.factor) %>% 
  mutate(across(Condition, factor, levels=c("Baseline", "Sitting","Light","Heavy")),
         Cond2 = ifelse(Condition == "Baseline", "Baseline", "Interaction"))

fsm <- fsm %>% 
  mutate_at(c("Task", "Condition", "Speaker"), as.factor) %>% 
  mutate(across(Condition, factor, levels=c("Baseline", "Sitting","Light","Heavy")),
         Cond2 = ifelse(Condition == "Baseline", "Baseline", "Interaction"))
```

# Confederate

## Read Speech

### Cycle Durations

```{r, include=FALSE}
dat <- brm %>%
  filter(Role=="Confederate", grepl("Read", Task))
```

```{r}
ggplot(dat, aes(Condition, cycleDur))+
  geom_boxplot()
```

* Difference only between Light and the other conditions, but no difference between Sitting and Heavy. So there is probably noise in the data and we can consider that there was no effect of condition on breathing cycle duration.

```{r}
dat$Condition <- relevel(dat$Condition, ref="Light")
summary(b1 <- lm(cycleDur ~ Condition, dat))

par(mfrow=c(2, 2))
hist(resid(b1))
qqnorm(resid(b1));qqline(resid(b1))
plot(fitted(b1), resid(b1))
```

### Inhalation Amplitude

```{r, include=FALSE}
dat <- brm %>%
  filter(Role=="Confederate", grepl("Read", Task))
```

```{r}
ggplot(dat, aes(Condition, inhalAmp))+
  geom_boxplot()
```

* Inhalation amplitude got higher with physical effort.

```{r}
dat$Condition <- relevel(dat$Condition, ref="Light")
summary(d1 <- lm(inhalAmp ~ Condition, dat))

par(mfrow=c(2, 2))
hist(resid(d1))
qqnorm(resid(d1));qqline(resid(d1))
plot(fitted(d1), resid(d1))
```

### F0

```{r, include=FALSE}
dat <- frb %>%
  filter(grepl("Read", Task), Speaker == "Confederate")
```

```{r}
ggplot(dat, aes(Condition, f0raw))+
  geom_boxplot()
```

* F0 got higher with physical effort.

```{r}
dat$Condition <- relevel(dat$Condition, ref="Light")
summary(f1 <- lm(f0raw ~ Condition, dat))

par(mfrow=c(2, 2))
hist(resid(f1))
qqnorm(resid(f1));qqline(resid(f1))
plot(fitted(f1), resid(f1))
```


## Free Speech

### Cycle Durations

```{r, include=FALSE}
dat <- brm %>%
  filter(Role=="Confederate", Task=="Free")
```

```{r}
ggplot(dat, aes(Condition, cycleDur))+
  geom_boxplot()
```

* Cycles are shorter during biking than during sitting. There's a trend of Heavy being shorter than Light, but it doesn't reach significance.

```{r}
dat$Condition <- relevel(dat$Condition, ref="Light")
summary(b1 <- lm(cycleDur ~ Condition, dat))

par(mfrow=c(2, 2))
hist(resid(b1))
qqnorm(resid(b1));qqline(resid(b1))
plot(fitted(b1), resid(b1))
```

### Inhalation Amplitude

```{r, include=FALSE}
dat <- brm %>%
  filter(Role=="Confederate", Task=="Free")
```

```{r}
ggplot(dat, aes(Condition, inhalAmp))+
  geom_boxplot()
```

* Inhalation amplitude gets larger with physical effort.

```{r}
dat$Condition <- relevel(dat$Condition, ref="Light")
summary(d1 <- lm(inhalAmp ~ Condition, dat))

par(mfrow=c(2, 2))
hist(resid(d1))
qqnorm(resid(d1));qqline(resid(d1))
plot(fitted(d1), resid(d1))
```

### F0

```{r, include=FALSE}
dat <- fsm %>%
  filter(Task == "Free", Speaker == "Confederate")
```

```{r}
ggplot(dat, aes(Condition, f0raw))+
  geom_boxplot()
```

* F0 increased with physical effort.

```{r}
dat$Condition <- relevel(dat$Condition, ref="Light")
summary(m1 <- lm(f0raw ~ Condition, dat))

par(mfrow=c(2, 2))
hist(resid(m1))
qqnorm(resid(m1));qqline(resid(m1))
plot(fitted(m1), resid(m1))
```


# Participants

## Solo Read Speech

### Cycle Durations

```{r, include=FALSE}
dat <- brm %>%
  filter(Role=="Participant", Task%in%c("ReadAlone", "ReadBaseline"))
```

```{r}
ggplot(dat, aes(Condition, cycleDur))+
  geom_boxplot()

ggplot(dat, aes(Cond2, cycleDur))+
  geom_boxplot()
```

* There is no effect of condition (or baseline vs interaction) on cycle duration during solo read speech.

```{r}
dat$Condition <- relevel(dat$Condition, ref="Light")
summary(b1 <- lmer(cycleDur ~ Condition + (1| Speaker), dat))
summary(b1a <- lmer(cycleDur ~ Cond2 + (1 + Cond2 | Speaker), dat))

par(mfrow=c(2, 2))
hist(resid(b1))
qqnorm(resid(b1));qqline(resid(b1))
plot(fitted(b1), resid(b1))

par(mfrow=c(2, 2))
hist(resid(b1a))
qqnorm(resid(b1a));qqline(resid(b1a))
plot(fitted(b1a), resid(b1a))
```

### Inhalation Amplitude

```{r, include=FALSE}
dat <- brm %>%
  filter(Role=="Participant", Task%in%c("ReadAlone", "ReadBaseline"))
```

```{r}
ggplot(dat, aes(Condition, inhalAmp))+
  geom_boxplot()

ggplot(dat, aes(Cond2, inhalAmp))+
  geom_boxplot()
```

* No effect of condition (or baseline vs. interaction) on inhalation amplitude during solo read speech.

```{r}
dat$Condition <- relevel(dat$Condition, ref="Light")

summary(a1 <- lmer(inhalAmp ~ Condition + (1 + Condition | Speaker), dat))
summary(a2 <- lmer(inhalAmp ~ Cond2 + (1 + Cond2 | Speaker), dat))

par(mfrow=c(2, 2))
hist(resid(a1))
qqnorm(resid(a1));qqline(resid(a1))
plot(fitted(a1), resid(a1))

par(mfrow=c(2, 2))
hist(resid(a2))
qqnorm(resid(a2));qqline(resid(a2))
plot(fitted(a2), resid(a2))
```

### F0

```{r, include=FALSE}
dat <- frb %>%
  filter(Role=="Participant", Task%in%c("ReadAlone", "ReadBaseline"))
```

```{r}
ggplot(dat, aes(Condition, f0raw))+
  geom_boxplot()

ggplot(dat, aes(Cond2, f0raw))+
  geom_boxplot()
```

* F0 was higher during interaction than during baseline.

```{r}
dat$Condition <- relevel(dat$Condition, ref="Light")

summary(f1 <- lmer(f0raw ~ Condition + (1 + Condition | Speaker), dat))
summary(f1a <- lmer(f0raw ~ Cond2 + (1 + Cond2 | Speaker), dat))

par(mfrow=c(2, 2))
hist(resid(f1))
qqnorm(resid(f1));qqline(resid(f1))
plot(fitted(f1), resid(f1))

par(mfrow=c(2, 2))
hist(resid(f1a))
qqnorm(resid(f1a));qqline(resid(f1a))
plot(fitted(f1a), resid(f1a))
```

## Synchronous Read Speech

### Cycle Durations

```{r, include=FALSE}
dat <- brm %>%
  filter(Role=="Participant", Task%in%c("ReadJoint", "ReadBaseline"))
```

```{r}
ggplot(dat, aes(Condition, cycleDur))+
  geom_boxplot()

ggplot(dat, aes(Cond2, cycleDur))+
  geom_boxplot()
```

* Cycle durations didn't change across conditions (nor baseline vs. interaction) during synchronous read speech.

```{r}
dat$Condition <- relevel(dat$Condition, ref="Light")

summary(b1 <- lmer(cycleDur ~ Condition + (1 | Speaker), dat))
summary(b1a <- lmer(cycleDur ~ Cond2 + (1 + Cond2 | Speaker), dat))

par(mfrow=c(2, 2))
hist(resid(b1))
qqnorm(resid(b1));qqline(resid(b1))
plot(fitted(b1), resid(b1))

par(mfrow=c(2, 2))
hist(resid(b1a))
qqnorm(resid(b1a));qqline(resid(b1a))
plot(fitted(b1a), resid(b1a))
```

### Inhalation Amplitude

```{r, include=FALSE}
dat <- brm %>%
  filter(Role=="Participant", Task%in%c("ReadJoint", "ReadBaseline"))
```

```{r}
ggplot(dat, aes(Condition, inhalAmp))+
  geom_boxplot()

ggplot(dat, aes(Cond2, inhalAmp))+
  geom_boxplot()
```

* Inhalation amplitude was lower during the interaction than during baseline.

```{r}
dat$Condition <- relevel(dat$Condition, ref="Light")

summary(a1 <- lmer(inhalAmp ~ Condition + (1 + Condition | Speaker), dat))
summary(a2 <- lmer(inhalAmp ~ Cond2 + (1 + Cond2 | Speaker), dat))

par(mfrow=c(2, 2))
hist(resid(a1))
qqnorm(resid(a1));qqline(resid(a1))
plot(fitted(a1), resid(a1))

par(mfrow=c(2, 2))
hist(resid(a2))
qqnorm(resid(a2));qqline(resid(a2))
plot(fitted(a2), resid(a2))
```

### F0

```{r, include=FALSE}
dat <- frb %>%
  filter(Role=="Participant", Task%in%c("ReadJoint", "ReadBaseline"))
```

```{r}
ggplot(dat, aes(Condition, f0raw))+
  geom_boxplot()

ggplot(dat, aes(Cond2, f0raw))+
  geom_boxplot()
```

* F0 tended to get higher with increase in confederate's physical effort (although no significant difference between Sitting and Light).

```{r}
dat$Condition <- relevel(dat$Condition, ref="Light")

summary(f1 <- lmer(f0raw ~ Condition + (1+Condition | Speaker), dat))

summary(f1a <- lmer(f0raw ~ Cond2 + (1+Cond2 | Speaker), dat))

par(mfrow=c(2, 2))
hist(resid(f1))
qqnorm(resid(f1));qqline(resid(f1))
plot(fitted(f1), resid(f1))

par(mfrow=c(2, 2))
hist(resid(f1a))
qqnorm(resid(f1a));qqline(resid(f1a))
plot(fitted(f1a), resid(f1a))
```

## Free Speech

### Cycle Durations

```{r, include=FALSE}
dat <- brm %>%
  filter(Role=="Participant", Task=="Free", act=="speaking")
```

```{r}
ggplot(dat, aes(Condition, cycleDur))+
  geom_boxplot()

ggplot(dat, aes(Cond2, cycleDur))+
  geom_boxplot()
```

* Cycle durations were shorter during Sitting and Baseline, and longer during Light and Heavy.

```{r}
dat$Condition <- relevel(dat$Condition, ref="Light")

summary(b1 <- lmer(cycleDur ~ Condition + (1| Speaker), dat))

par(mfrow=c(2, 2))
hist(resid(b1))
qqnorm(resid(b1));qqline(resid(b1))
plot(fitted(b1), resid(b1))
```

### Inhalation Amplitude

```{r, include=FALSE}
dat <- brm %>%
  filter(Role=="Participant", Task=="Free", act=="speaking")
```

```{r}
ggplot(dat, aes(Condition, inhalAmp))+
  geom_boxplot()

ggplot(dat, aes(Cond2, inhalAmp))+
  geom_boxplot()

```

* No effect of conditon (nor baseline vs interaction) on inhalation amplitude during free speech.

```{r}
dat$Condition <- relevel(dat$Condition, ref="Light")

summary(a1 <- lmer(inhalAmp ~ Condition + (1 + Condition| Speaker), dat))
summary(a2 <- lmer(inhalAmp ~ Cond2 + (1 + Cond2| Speaker), dat))

par(mfrow=c(2, 2))
hist(resid(a1))
qqnorm(resid(a1));qqline(resid(a1))
plot(fitted(a1), resid(a1))

par(mfrow=c(2, 2))
hist(resid(a2))
qqnorm(resid(a2));qqline(resid(a2))
plot(fitted(a2), resid(a2))
```

### F0

```{r, include=FALSE}
dat <- fsm %>%
  filter(Role=="Participant", Task=="Free")
```

```{r}
ggplot(dat, aes(Condition, f0raw))+
  geom_boxplot()

ggplot(dat, aes(Cond2, f0raw))+
  geom_boxplot()
```

* No effect of condition (or baseline vs interaction) on f0 during free speech.

```{r}
dat$Condition <- relevel(dat$Condition, ref="Light")

summary(m1 <- lmer(f0raw ~ Condition + (1 + Condition | Speaker), dat))
summary(m2 <- lmer(f0raw ~ Cond2 + (1 + Cond2 | Speaker), dat))

par(mfrow=c(2, 2))
plot(fitted(m1), residuals(m1))
qqnorm(resid(m1));qqline(resid(m1)) # residuals ok I think
hist(resid(m1))

par(mfrow=c(2, 2))
plot(fitted(m2), residuals(m2))
qqnorm(resid(m2));qqline(resid(m2)) # residuals ok I think
hist(resid(m2))
```

## Watching vs. Listening

### Cycle Durations

```{r, include=FALSE}
dat <- brm %>%
  filter(Role=="Participant", act %in% c("watching", "listening"))
```

```{r}
ggplot(dat %>% filter(cycleDur < 10), aes(Condition, cycleDur))+
  geom_boxplot()+
  facet_wrap(~act)

ggplot(dat %>% filter(cycleDur < 10), aes(act, cycleDur))+
  geom_boxplot()
```

* Cycle durations are shorter during listening than baseline.

* There was no effect of condition on cycle duration within Watching nor within Listening.

```{r}
summary(c1 <- lmer(cycleDur ~ act + (1+ act |Speaker), dat))

summary(c1.1a <- lmer(cycleDur ~ Condition + (1 + Condition |Speaker), dat %>% filter(act=="listening"))) # random slope causes the model not to converge
summary(c1.2 <- lmer(cycleDur ~ Condition + (1 + Condition |Speaker), dat %>% filter(act=="watching"))) # random slope causes the model not to converge

par(mfrow=c(2, 2))
hist(resid(c1))
qqnorm(resid(c1));qqline(resid(c1))
plot(fitted(c1), resid(c1))
```

### Inhalation Amplitude

```{r, include=FALSE}
dat <- brm %>%
  filter(Role=="Participant", act %in% c("watching", "listening"))
```

```{r}
ggplot(dat, aes(Condition, inhalAmp))+
  geom_boxplot()+
  facet_wrap(~act)

ggplot(dat, aes(act, inhalAmp))+
  geom_boxplot()
```

* Inhalation amplitude was lower during Listening than Watching.

* During Listening, there was no condition effect.

* During Watching, there is a difference between Sitting and Heavy (with Light in the middle). So there seems to be a small condition effect.

```{r}
summary(a3 <- lmer(inhalAmp ~ act + (1 + act | Speaker), dat))
summary(a4 <- lmer(inhalAmp ~ Condition + (1 + Condition| Speaker), dat %>% filter(act=="watching")))
summary(a5 <- lmer(inhalAmp ~ Condition + (1 +Condition | Speaker), dat %>% filter(act=="listening")))

par(mfrow=c(2, 2))
hist(resid(a3))
qqnorm(resid(a3));qqline(resid(a3))
plot(fitted(a3), resid(a3))
```