---
title: "Analysis-Defense"
author: "Tom Offrede"
date: "2024-04-29"
output:
  html_document:
    toc: true
    toc_float: 
      collapsed: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
Sys.setenv(LANG="en")
```

```{r, include=FALSE}
library(lmerTest)
library(tidyverse)
library(broom.mixed)
library(ggpubr)
library(rPraat)
`%!in%` <- Negate(`%in%`)

folder <- "C:/Users/offredet/Documents/1HU/ExperimentBreathing/Data/DataForAnalysis/AllData/"
folderInd <- "C:/Users/offredet/Documents/1HU/ExperimentBreathing/Defense/" # to save plots with individual effects
folderBreath <- "C:/Users/offredet/Documents/1HU/ExperimentBreathing/Data/DataForAnalysis/PeaksValleys/"

orderCond <- c("Sitting", "Light", "Heavy")
orderCondBase <- c("Baseline", orderCond)
orderAct <- c("watching", "listening")

load(paste0(folder, "DataBreathing2.RData"))
load(paste0(folder, "DataConfBaselineBreathing.RData"))

brm <- brm2 |> rename(speaker = Speaker, condition = Condition) |> 
  mutate(Ifraction = inhalDur / cycleDur,
         act2 = ifelse(grepl("Read", act), "read", ifelse(grepl("speak", act), "speak", as.character(act)))) |> # I-fraction is defined by Wilder & Baker 1974 as "the ratio of the length of inspiration to the total duration of the respiratory cycle"
  filter(speaker!="Confederate") |> 
  mutate(text = case_when(
    List == "1" & substr(file, 1, 2) == "SR" ~ "Hirsch",
    List == "1" & substr(file, 1, 2) == "LR" ~ "Schwalbe",
    List == "1" & substr(file, 1, 2) == "HR" ~ "Pferd",
    List == "2" & substr(file, 1, 2) == "SR" ~ "Pferd",
    List == "2" & substr(file, 1, 2) == "LR" ~ "Hirsch",
    List == "2" & substr(file, 1, 2) == "HR" ~ "Schwalbe",
    List == "3" & substr(file, 1, 2) == "SR" ~ "Schwalbe",
    List == "3" & substr(file, 1, 2) == "LR" ~ "Pferd",
    List == "3" & substr(file, 1, 2) == "HR" ~ "Hirsch",
    List == "4" & substr(file, 1, 2) == "SR" ~ "Hirsch",
    List == "4" & substr(file, 1, 2) == "LR" ~ "Pferd",
    List == "4" & substr(file, 1, 2) == "HR" ~ "Schwalbe",
    List == "5" & substr(file, 1, 2) == "SR" ~ "Hirsch",
    List == "5" & substr(file, 1, 2) == "LR" ~ "Schwalbe",
    List == "5" & substr(file, 1, 2) == "HR" ~ "Pferd",
    List == "6" & substr(file, 1, 2) == "SR" ~ "Schwalbe",
    List == "6" & substr(file, 1, 2) == "LR" ~ "Hirsch",
    List == "6" & substr(file, 1, 2) == "HR" ~ "Pferd"
  )) |> 
  mutate_at(c("text"), as.factor)
brm <- droplevels(brm)
brc0 <- brm2 |> 
  rename(speaker = Speaker, condition = Condition) |> 
  filter(speaker=="Confederate") |> 
  select(names(bc))
brc <- rbind(brc0, bc) |> 
  mutate(Ifraction = inhalDur / cycleDur,
         text = case_when(
           grepl("Hirs", file) ~ "Hirsch",
           grepl("Pfer", file) ~ "Pferd",
           grepl("Schw", file) ~ "Schwalbe"
         ))
brc <- droplevels(brc)
remove(brm2)
remove(brc0)
remove(bc)

# speech data
load(paste0(folder, "DataSpeech.RData"))
load(paste0(folder, "DataReadSpeech.RData"))

ds <- rbind(frb |> 
              select(Speaker, file, f0raw, Task, IPU, Condition, Role) |> 
              relocate(Speaker, file, IPU, f0raw, Condition, Task, Role),
            fsm |> 
              select(Speaker, file, IPU, f0raw, Condition, Task) |> 
              mutate(Role = ifelse(Speaker=="Confederate", "Confederate", "Participant")) |> 
              relocate(Speaker, file, IPU, f0raw, Condition, Task, Role) # make sure they're both in the same order
) |> 
  mutate(Task = ifelse(as.character(Task) == "ReadBaseline", "ReadAlone", as.character(Task)),
         Cond2 = ifelse(Condition == "Baseline", "Baseline", "WithConf")) |> 
  mutate_at(c("Cond2"), as.factor)
```

# Participants vs. Confederate

```{r}
# create dataset for cross correlation of confederate's and participants' breathing

files0 <- list.files(folderBreath, ".TextGrid")
filesT <- files0[substr(files0, 2,2) == "-" & grepl("joint", files0)]

# files0 <- list.files(folderBreath, ".wav")
# filesW <- files0[substr(files0, 2,2) == "-" & grepl("joint", files0)]
# files <- data.frame(wav = filesW,
#                     tg = filesT)

dc <- data.frame(matrix(nrow=0, ncol=9))
names(dc) <- c("speaker", "file", "text", "condition", "breathCycle", "onset", "peak", "offset", "cycleDur")

for(f in filesT){
  tg <- tg.read(paste0(folderBreath, f), encoding = detectEncoding(paste0(folderBreath, f)))
  for(t in 1:(tg.getNumberOfPoints(tg, 2)-1)){
    dc[nrow(dc)+1,] <- c("Confederate",
                         f,
                         strsplit(strsplit(f, "_")[[1]][1], "-")[[1]][2],
                         case_when(
                           substr(f, 1, 1) == "S" ~ "Sitting",
                           substr(f, 1, 1) == "L" ~ "Light",
                           substr(f, 1, 1) == "H" ~ "Heavy"),
                         paste0("cycle", t),
                         as.numeric(tg.getPointTime(tg, 2, t)),
                         as.numeric(tg.getPointTime(tg, 1, t)),
                         as.numeric(tg.getPointTime(tg, 2, t+1)), # next valley
                         (as.numeric(tg.getPointTime(tg, 2, t+1))) - (as.numeric(tg.getPointTime(tg, 2, t)))
    )
  }
}
dc <- dc |> 
  mutate_at(c("speaker", "file", "text", "condition", "breathCycle"), as.factor) |> 
  mutate_at(c("onset", "peak", "offset", "cycleDur"), as.numeric) |> 
  group_by(file) |> 
               mutate(offset = offset - min(onset),
                      onset = onset - min(onset)) |> 
               ungroup()

ccb <- rbind(brm |> 
               filter(act=="ReadJoint") |> 
               select(speaker, file, text, breathCycle, onset, peak, offset, cycleDur, condition) |> 
               relocate(speaker, file, text, condition, breathCycle, onset, peak, offset, cycleDur) |> 
               group_by(file) |> 
               mutate(offset = offset - min(onset),
                      onset = onset - min(onset)) |> 
               ungroup(),
             dc)
```

Check out `C:/Users/offredet/Documents/1HU/ExperimentBreathing/Data/DataForAnalysis/PeaksValleys/` for the plots created below.
I had originally also created these plots for cycle onset, but they're almost exactly like the ones for peak, so I deleted those.

```{r}
files0 <- list.files(folderBreath, "wav$")
filesP <- files0[grepl("joint", files0) & substr(files0, 2, 2) =="R"]
filesC <- files0[grepl("joint", files0) & substr(files0, 2, 2) =="-"]

cross <- data.frame(matrix(nrow=0, ncol=4))
names(cross) <- c("speaker", "condition", "text", "lag0ccf")

for(s in unique(ccb$speaker[ccb$speaker!="Confederate"])){
  for(c in c("Sitting", "Light", "Heavy")){
    filenameP <- filesP[grepl(paste0(substr(c, 1, 1), "R-", s), filesP)]
    filenameC <- filesC[grepl(paste0(substr(c, 1, 1), "-", unique(ccb$text[ccb$speaker==s & ccb$condition==c])), filesC)]

    bp <- tuneR::readWave(paste0(folderBreath, filenameP))
    bw <- data.frame(w = bp@left,
                     time = seq(0, (nrow(data.frame(bp@left))/2), length.out=nrow(data.frame(bp@left))))
    pb <- ggplot(bw, aes(time, w))+
      geom_line(linewidth=1)+
      labs(y=paste("Participant:", s))

    bc <- tuneR::readWave(paste0(folderBreath, filenameC))
    cw <- data.frame(w = bc@left,
                     time = seq(0, (nrow(data.frame(bc@left))/2), length.out=nrow(data.frame(bc@left))))
    cb <- ggplot(cw, aes(time, w))+
      geom_line(linewidth=1)+
      labs(y="Confederate")

    results0 <- print(ccf(ccb$peak[ccb$speaker==s & ccb$condition==c],
                          ccb$peak[ccb$speaker=="Confederate" & ccb$text==ccb$text[ccb$speaker==s & ccb$condition==c] & ccb$condition==c],
                          plot=FALSE))
    results <- data.frame(lag=results0$lag, acf=results0$acf)
    sig_limit <- 2 / sqrt(results0$n.used)

    (ccf_plot <- ggplot(results, aes(lag, acf)) +
        geom_bar(stat = "identity", width=0.1) +
        geom_hline(yintercept = c(sig_limit, -sig_limit), linetype = "dashed", color = "blue") +
        labs(title = "PEAK Cross-Correlation", x = "Lag", y = "ACF"))

    ggarrange(cb, pb, ccf_plot, ncol=1,
              bottom=text_grob(paste("-4:", results$acf[results$lag==-4], " -3:", results$acf[results$lag==-3], " -2:", results$acf[results$lag==-2], " -1:", results$acf[results$lag==-1], " 0:", results$acf[results$lag==0], " 1:", results$acf[results$lag==1], " 2:", results$acf[results$lag==2], " 3:", results$acf[results$lag==3], " 4:", results$acf[results$lag==4])))

    ggsave(paste0(folderInd, s, "-", c, "-Peak.png"), width=3500, height=5000, units="px", dpi=500)
    
    cross[nrow(cross)+1,] <- c(s, c, ccb$text[ccb$speaker==s & ccb$condition==c], results$acf[results$lag==0])

  }
}
cross <- cross |> 
  mutate_at(c("speaker", "condition", "text"), as.factor) |> 
  mutate_at(c("lag0ccf"), as.numeric)
```

```{r}
ggplot(cross, aes(condition, lag0ccf))+
  geom_boxplot()

summary(lm(lag0ccf ~ condition, cross)) # adding random effects isn't working
```


```{r}
b <- tuneR::readWave(paste0(folderBreath, "SR-ATN010_SUM_200_joint_breath.wav"))
bw <- data.frame(w = b@left,
                 time = seq(0, (nrow(data.frame(b@left))/2), length.out=nrow(data.frame(b@left))))
pb <- ggplot(bw, aes(time, w))+
  geom_line(linewidth=1)+
  labs(y="Participant")

c <- tuneR::readWave(paste0(folderBreath, "S-Pferd_THORAX_joint_200.wav"))
cw <- data.frame(w = c@left,
                 time = seq(0, (nrow(data.frame(c@left))/2), length.out=nrow(data.frame(c@left))))
cb <- ggplot(cw, aes(time, w))+
  geom_line(linewidth=1)+
  labs(y="Confederate")

ggarrange(cb, pb, nrow=2)



```

# Watching vs. Listening

## Cycle duration

Breath cycles are shorter during listening, i.e. faster respiration rate. This is in line with previous literature (listening breathing is somewhere between quiet and speech breathing).

```{r}
ggplot(brm |> filter(act%in%c("watching", "listening"), cycleDur<7.5), aes(act, cycleDur))+
  geom_boxplot()+
  scale_x_discrete(limits=orderAct)

summary(lmer(cycleDur ~ act + (1+act|speaker), brm |> filter(act%in%c("watching", "listening"))))
```

**No condition effect during listening**

```{r}
ggplot(brm |> filter(act%in%c("listening"), cycleDur<7.5), aes(condition, cycleDur))+
  geom_boxplot()

summary(lmer(cycleDur ~ condition + (1|speaker), brm |> filter(act%in%c("listening"))))
```

## Inhalation amplitude

Inhalation amplitude is much lower while listening than watching

```{r}
ggplot(brm |> filter(#!is.na(act)
  act%in%c("watching", "listening")
), aes(act, inhalAmp))+
  geom_boxplot()+
  scale_x_discrete(limits=orderAct)

summary(lmer(inhalAmp ~ act + (1+act|speaker), brm |> filter(act%in%c("watching", "listening"))))
```

**No condition effect during listening**

```{r}
ggplot(brm |> filter(act%in%c("listening")), aes(condition, inhalAmp))+
  geom_boxplot()

summary(lmer(inhalAmp ~ condition + (1+condition|speaker), brm |> filter(act%in%c("listening"))))
```

## I-fraction

```{r}
ggplot(brm |> filter(#!is.na(act)
  act%in%c("watching", "listening")
), aes(act, Ifraction))+
  geom_boxplot()

summary(lmer(Ifraction ~ act + (1+act|speaker), brm |> filter(act%in%c("watching", "listening"))))
```

**No condition effect for listening nor watching**

```{r}
ggplot(brm |> filter(#!is.na(act)
  act%in%c("watching", "listening")
), aes(condition, Ifraction))+
  geom_boxplot()+
  facet_wrap(~act)

summary(lmer(Ifraction ~ condition + (1|speaker), brm |> filter(act%in%c("listening"))))

summary(lmer(Ifraction ~ condition + (1|speaker), brm |> filter(act%in%c("watching"))))
```

# Confederate

### f0

* Similar condition effect in Read and Free

* Read has higher f0, but this is accounted for only by the Heavy condition. Sitting and Light had similar f0 in both Read and Free

```{r}
ggplot(ds |> filter(Role=="Confederate"), aes(Condition, f0raw))+
  geom_boxplot()+
  facet_wrap(~Task)

summary(m <- lm(f0raw ~ Condition, ds |> filter(Task=="ReadAlone", Role=="Confederate")))

summary(m <- lm(f0raw ~ Condition, ds |> filter(Task=="Free", Role=="Confederate")))

ggplot(ds |> filter(Role=="Participant", Task%in%c("Free", "ReadAlone")), aes(Task, f0raw))+
  geom_boxplot()

summary(m <- lm(f0raw ~ Task * Condition, ds |> filter(Task%in%c("Free", "ReadAlone"), Role=="Confederate")))

summary(emmeans::emmeans(m, pairwise ~ Task|Condition))

```

## Cycle duration

In read speech, we had found a weird pattern for cycle duration (reported in paper).

In spontaneous speech, the expected pattern is found: faster breathing with higher effort (although, no difference between Light and Heavy).

Overall, breathing also tends to be faster during spontaneous than read speech, but only slightly so (p = 0.042).

```{r}
brc$condition <- relevel(brc$condition, ref="Light")

ggplot(brc|> filter(act!="watching"), aes(condition, cycleDur))+
  geom_boxplot()+
  facet_wrap(~act)+
  scale_x_discrete(limits=orderCond)

summary(lm(cycleDur ~ condition, brc |> filter(act=="ReadAlone")))
summary(lm(cycleDur ~ condition, brc |> filter(act=="speaking")))

ggplot(brc|> filter(act!="watching"), aes(act, cycleDur))+
  geom_boxplot()

summary(lm(cycleDur ~ act, brc |> filter(act!="watching")))
```

## Inhalation amplitude

Inhalation amplitude increased more with condition during read speech than with spontaneous speech (during sitting there was no difference between read and spontaneous). In other words: overall, amplitude is larger in read than spontaneous speech; and this is accounted for by the larger effect of condition on amplitude in read speech.

```{r}
ggplot(brc|> filter(act!="watching"), aes(condition, inhalAmp))+
  geom_boxplot()+
  facet_wrap(~act)+
  scale_x_discrete(limits=orderCond)

summary(lm(inhalAmp ~ condition, brc |> filter(act=="ReadAlone")))
summary(lm(inhalAmp ~ condition, brc |> filter(act=="speaking")))

summary(m <- lm(inhalAmp ~ condition * act, brc |> filter(act%in%c("speaking", "ReadAlone"))))
summary(emmeans::emmeans(m, pairwise ~ act|condition))

ggplot(brc|> filter(act!="watching"), aes(act, inhalAmp))+
  geom_boxplot()

summary(lm(inhalAmp ~ act, brc |> filter(act!="watching")))
```


## I-fraction

**New thoughts about I-fraction: in this context, I'm not so interested in this variable, because what we could learn from it we could probably learn with more accuracy from inhalation duration**

In both read and free speech, I-fraction generally increases with increase in effort (although no difference between Light and Heavy during spontaneous speech, as was the case for cycle duration; and no difference between Sitting and Light during read speech).

While biking, I-fraction is higher in spontaneous speech.

(During sitting, I-fraction was slightly higher during read speech.)

Remember: higher I-fraction = more balanced inhalation / exhalation durations.

```{r}
ggplot(brc|> filter(act!="watching"), aes(condition, Ifraction))+
  geom_boxplot()+
  facet_wrap(~act)+
  scale_x_discrete(limits=orderCond)

m <- lm(Ifraction ~ condition * act, brc |> filter(act!="watching"))
summary(emmeans::emmeans(m, pairwise ~ condition|act))
summary(emmeans::emmeans(m, pairwise ~ act|condition))


ggplot(brc|> filter(act!="watching"), aes(act, Ifraction))+
  geom_boxplot()

summary(lm(Ifraction ~ act, brc |> filter(act!="watching")))
```


# Participants: exploring read vs spontaneous speech

## General comparisons

### f0

* Read alone: no condition difference (in paper); but f0 higher during experiment than at baseline

* Free: no differences between conditions or between baseline vs during experiment

* Free vs Read: higher f0 in read speech

```{r}
ggplot(ds |> filter(Role=="Participant", Task%in%c("Free", "ReadAlone")), aes(Condition, f0raw))+
  geom_boxplot()+
  facet_wrap(~Task)

summary(m <- lmer(f0raw ~ Condition + (1 + Condition | Speaker), ds |> filter(Task=="ReadAlone", Role=="Participant", Condition != "Baseline")))
summary(m <- lmer(f0raw ~ Cond2 + (1 + Cond2 | Speaker), ds |> filter(Task=="ReadAlone", Role=="Participant")))

summary(m <- lmer(f0raw ~ Condition + (1 + Condition | Speaker), ds |> filter(Task=="Free", Role=="Participant", Condition != "Baseline")))
summary(m <- lmer(f0raw ~ Cond2 + (1 + Cond2 | Speaker), ds |> filter(Task=="Free", Role=="Participant")))

ggplot(ds |> filter(Role=="Participant", Task%in%c("Free", "ReadAlone")), aes(Task, f0raw))+
  geom_boxplot()

summary(m <- lmer(f0raw ~ Task + (1 + Task | Speaker), ds |> filter(Task%in%c("Free", "ReadAlone"), Role=="Participant", Condition!="Baseline")))
```


### Cycle duration

* Free speech: slower breathing rate during biking conditions, as compared to sitting (and sitting was same as baseline)

* Read speech: no condition effect in either alone nor joint. Also no overall difference between baseline, alone and joint.

* Free vs read: overall, longer cycles (slower rate) during free, than read, speech

```{r}
brm$condition <- relevel(brm$condition, ref="Light")

ggplot(brm|> filter(act%!in%c("watching", "listening", "ReadBaseline", "speakingBaseline")), aes(condition, cycleDur))+
  geom_boxplot()+
  facet_wrap(~act)+
  scale_x_discrete(limits=orderCond)

summary(lmer(cycleDur ~ condition + (1|speaker), brm |> filter(act=="ReadAlone")))
summary(lmer(cycleDur ~ condition + (1|speaker), brm |> filter(act=="ReadJoint")))
summary(lmer(cycleDur ~ condition + (1|speaker), brm |> filter(act=="speakingConditions")))

ggplot(brm|> filter(act%!in%c("watching", "listening")), aes(act, cycleDur))+
  geom_boxplot()

summary(lmer(cycleDur ~ act + (1|speaker), brm |> filter(grepl("Read|speak", act))))
```


### Inhalation amplitude

* Free speech: no robust condition differences.

* Read speech: **alone**: no condition difference. **joint**: light with higher amplitude (already in paper).

* Free vs read: higher amplitudes at baseline and alone, and lower at joint and free. (overall, also a decreasing pattern)

```{r}
brm$condition <- relevel(brm$condition, ref="Light")
brm$act <- relevel(brm$act, ref="ReadJoint")

ggplot(brm|> filter(act%!in%c("watching", "listening", "ReadBaseline", "speakingBaseline")), aes(condition, inhalAmp))+
  geom_boxplot()+
  facet_wrap(~act)+
  scale_x_discrete(limits=orderCond)

summary(lmer(inhalAmp ~ condition + (1+condition|speaker), brm |> filter(act=="ReadAlone")))
summary(lmer(inhalAmp ~ condition + (1|speaker), brm |> filter(act=="ReadJoint")))
summary(lmer(inhalAmp ~ condition + (1+condition|speaker), brm |> filter(act=="speakingConditions")))

ggplot(brm|> filter(act%!in%c("watching", "listening")), aes(act, inhalAmp))+
  geom_boxplot()+
  scale_x_discrete(limits=c("ReadBaseline", "ReadAlone", "ReadJoint", "speakingBaseline", "speakingConditions"))

summary(lmer(inhalAmp ~ act + (1+act|speaker), brm |> filter(grepl("Read|speak", act))))
```


### I-fraction

* Free speech: Sitting has larger I-fraction than Heavy (Light is not different from either)

* Read speech: **alone**: no condition difference. **joint**: higher I-fraction with higher effort (but no difference between light and heavy)

* Free vs read: no overall differences, just a few specific ones which may have been due to sample size (e.g. I-fraction larger in speakingConditions than speakingBaseline, and in ReadJoint than in all other categories)


```{r}
brm$condition <- relevel(brm$condition, ref="Sitting")

ggplot(brm|> filter(act%!in%c("watching", "listening", "ReadBaseline", "speakingBaseline")), aes(condition, Ifraction))+
  geom_boxplot()+
  facet_wrap(~act)+
  scale_x_discrete(limits=orderCond)

summary(lmer(Ifraction ~ condition + (1|speaker), brm |> filter(act=="ReadAlone")))
summary(lmer(Ifraction ~ condition + (1|speaker), brm |> filter(act=="ReadJoint")))
summary(lmer(Ifraction ~ condition + (1|speaker), brm |> filter(act=="speakingConditions")))

ggplot(brm|> filter(act%!in%c("watching", "listening")), aes(act, Ifraction))+
  geom_boxplot()+
  scale_x_discrete(limits=c("ReadBaseline", "ReadAlone", "ReadJoint", "speakingBaseline", "speakingConditions"))

brm$act <- relevel(brm$act, ref="ReadAlone")
summary(lmer(Ifraction ~ act + (1|speaker), brm |> filter(grepl("Read|speak", act))))

ggplot(brm|> filter(act2%in%c("read", "speak")), aes(act2, Ifraction))+
  geom_boxplot()

summary(lmer(Ifraction ~ act2 + (1+act2|speaker), brm |> filter(grepl("read|speak", act2))))
```

## Comparison: speakBaseline vs ReadBaseline

### f0

* Higher f0 in read speech (as for all the conditions)

```{r}
ggplot(ds |> filter(Role=="Participant", Task%in%c("Free", "ReadAlone"), Condition =="Baseline"), aes(Task, f0raw))+
  geom_boxplot()

summary(m <- lmer(f0raw ~ Task + (1 + Task | Speaker), ds |> filter(Task%in%c("Free", "ReadAlone"), Role=="Participant", Condition=="Baseline")))
```

### Cycle duration

* Longer cycles (slower breathing rate) during free speech

```{r}
ggplot(brm |> filter(act%in%c("speakingBaseline", "ReadBaseline")), aes(act, cycleDur))+
  geom_boxplot()

summary(lmer(cycleDur ~ act + (1+act|speaker), brm |> filter(act%in%c("speakingBaseline", "ReadBaseline"))))
```

### Inhalation amplitude

* Lower inhalation amplitudes in free speech

```{r}
ggplot(brm |> filter(act%in%c("speakingBaseline", "ReadBaseline")), aes(act, inhalAmp))+
  geom_boxplot()

summary(lmer(inhalAmp ~ act + (1+act|speaker), brm |> filter(act%in%c("speakingBaseline", "ReadBaseline"))))
```

### I-fraction

* No difference in I-fraction

Interesting, given the robust differences in cycle duration and inhalation amplitude!

```{r}
ggplot(brm |> filter(act%in%c("speakingBaseline", "ReadBaseline")), aes(act, Ifraction))+
  geom_boxplot()

summary(lmer(Ifraction ~ act + (1+act|speaker), brm |> filter(act%in%c("speakingBaseline", "ReadBaseline"))))
```

## Comparison: speakConditions vs ReadAlone

### f0

* Higher f0 in read speech (as for baseline)

```{r}
ggplot(ds |> filter(Role=="Participant", Task%in%c("Free", "ReadAlone"), Condition!="Baseline"), aes(Task, f0raw))+
  geom_boxplot()

summary(m <- lmer(f0raw ~ Task + (1 + Task | Speaker), ds |> filter(Task%in%c("Free", "ReadAlone"), Role=="Participant", Condition!="Baseline")))
```

### Cycle duration

* Longer cycles (slower breathing rate) during free speech (same as baseline)

```{r}
ggplot(brm |> filter(act%in%c("speakingConditions", "ReadAlone")), aes(#act, cycleDur))+
  geom_boxplot()
  
  summary(lmer(cycleDur ~ act + (1+act|speaker), brm |> filter(act%in%c("speakingConditions", "ReadAlone"))))
  ```
  
  * Read: no condition difference
  
  * Free: sitting -- shorter cycle duration (faster rate) than biking conditions
  
  ```{r}
  ggplot(brm |> filter(act%in%c("speakingConditions", "ReadAlone")), aes(condition, cycleDur))+
    geom_boxplot()+
    facet_wrap(~act)
  
  summary(lmer(cycleDur ~ condition + (1|speaker), brm |> filter(act%in%c("ReadAlone"))))
  
  brm$condition <- relevel(brm$condition, ref="Light")
  summary(lmer(cycleDur ~ condition + (1|speaker), brm |> filter(act%in%c("speakingConditions"))))
  ```
  
  ### Inhalation amplitude
  
  * Tendency to lower inhalation amplitudes in free speech (like baseline), but difference not significant
  
  ```{r}
  ggplot(brm |> filter(act%in%c("speakingConditions", "ReadAlone")), aes(act, inhalAmp))+
    geom_boxplot()
  
  summary(lmer(inhalAmp ~ act + (1+act|speaker), brm |> filter(act%in%c("speakingConditions", "ReadAlone"))))
  ```
  
  * Read: no condition difference
  
  * Free: no condition difference
  
  ```{r}
  ggplot(brm |> filter(act%in%c("speakingConditions", "ReadAlone")), aes(condition, inhalAmp))+
    geom_boxplot()+
    facet_wrap(~act)
  
  summary(lmer(inhalAmp ~ condition + (1|speaker), brm |> filter(act%in%c("ReadAlone"))))
  
  brm$condition <- relevel(brm$condition, ref="Light")
  summary(lmer(inhalAmp ~ condition + (1|speaker), brm |> filter(act%in%c("speakingConditions"))))
  ```
  
  ### I-fraction
  
  * No difference in I-fraction (same as baseline)
  
  ```{r}
  ggplot(brm |> filter(act%in%c("speakingConditions", "ReadAlone")), aes(act, Ifraction))+
    geom_boxplot()
  
  summary(lmer(Ifraction ~ act + (1+act|speaker), brm |> filter(act%in%c("speakingConditions", "ReadAlone"))))
  ```
  
  * Read: no condition difference
  
  * Free: no condition difference
  
  ```{r}
  ggplot(brm |> filter(act%in%c("speakingConditions", "ReadAlone")), aes(condition, Ifraction))+
    geom_boxplot()+
    facet_wrap(~act)
  
  summary(lmer(Ifraction ~ condition + (1|speaker), brm |> filter(act%in%c("ReadAlone"))))
  
  brm$condition <- relevel(brm$condition, ref="Light")
  summary(lmer(Ifraction ~ condition + (1|speaker), brm |> filter(act%in%c("speakingConditions"))))
  ```
  
  ## Since (almost) no difference between conditions, conflate Baseline and Alone
  
  ```{r}
  dat <- brm |> 
    mutate_at(c("act"), as.character) |> 
    mutate(act = ifelse(act %in% c("speakingBaseline", "speakingConditions"), "free", ifelse(act %in% c("ReadBaseline", "ReadAlone"), "read", NA))) |> 
    filter(!is.na(act)) |> 
    mutate_at(c("act"), as.factor)
  ```
  
  ### Cycle duration
  
  * Longer cycles (slower breathing rate) during free speech
  
  ```{r}
  ggplot(dat, aes(act, cycleDur))+
    geom_boxplot()
  
  ggplot(dat, aes(act, cycleDur))+
    geom_boxplot()+
    facet_wrap(~speaker)
  
  summary(lmer(cycleDur ~ act + (1+act|speaker), dat))
  ```
  
  ### Inhalation amplitude
  
  * Lower inhalation amplitudes in free speech
  
  ```{r}
  ggplot(dat, aes(act, inhalAmp))+
    geom_boxplot()
  
  ggplot(dat, aes(act, inhalAmp))+
    geom_boxplot()+
    facet_wrap(~speaker)
  
  summary(lmer(inhalAmp ~ act + (1+act|speaker), dat))
  ```
  
  ### I-fraction
  
  * No difference in I-fraction
  
  Makes sense, given that both inhalation AND exhalation durations were longer in free speech. So participants divide their breath cycles equally regardless of the type of speech.
  
  ```{r}
  ggplot(dat, aes(act, Ifraction))+
    geom_boxplot()
  
  ggplot(dat, aes(act, Ifraction))+
    geom_boxplot()+
    facet_wrap(~speaker)
  
  summary(lmer(Ifraction ~ act + (1+act|speaker), dat))
  ```
  
  