---
title: "Analysis - SMC Paper Structured"
author: "Tom Offrede"
date: "2023-01-27"
output:
  html_document:
    toc: true
    toc_float: 
      collapsed: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, include=FALSE}
library(lme4)
library(tidyverse)
library(broom)

folder <- "C:/Users/offredet/Documents/1HU/ExperimentBreathing/Data/DataForAnalysis/AllData/"
folderInd <- "C:/Users/offredet/Documents/1HU/ExperimentBreathing/Data/IndividualEffects/" # to save plots with individual effects

orderCond <- c("Sitting", "Light", "Heavy")
orderCondBase <- c("Baseline", orderCond)
orderAct <- c("watching", "listening")

load(paste0(folder, "DataBreathing.RData"))
load(paste0(folder, "DataReadSpeech.RData"))
load(paste0(folder, "DataSpeech.RData"))
```

# Confederate

## Read Speech

### Cycle Durations

```{r}
dat <- brm %>%
  filter(Role=="Confederate", grepl("Read", Task))
```

```{r}
ggplot(dat, aes(Condition, cycleDur))+
  geom_boxplot()
```

* Difference only between Light and the other conditions, but no difference between Sitting and Heavy. So there is probably noise in the data and we can consider that there was no effect of condition on breathing cycle duration.

```{r}
dat$Condition <- relevel(dat$Condition, ref="Light")
summary(b1 <- lm(cycleDur ~ Condition, dat))

par(mfrow=c(2, 2))
hist(resid(b1))
qqnorm(resid(b1));qqline(resid(b1))
plot(fitted(b1), resid(b1))
```

### Inhalation Amplitude

```{r}
dat <- brm %>%
  filter(Role=="Confederate", grepl("Read", Task))
```

```{r}
ggplot(dat, aes(Condition, inhalAmp))+
  geom_boxplot()
```

* Inhalation amplitude got higher with physical effort.

```{r}
dat$Condition <- relevel(dat$Condition, ref="Light")
summary(d1 <- lm(inhalAmp ~ Condition, dat))

par(mfrow=c(2, 2))
hist(resid(d1))
qqnorm(resid(d1));qqline(resid(d1))
plot(fitted(d1), resid(d1))
```

### F0

```{r}
dat <- frb %>%
  filter(grepl("Read", Task), Speaker == "Confederate")
```

```{r}
ggplot(dat, aes(Condition, f0raw))+
  geom_boxplot()
```

* F0 got higher with physical effort.

```{r}
dat$Condition <- relevel(dat$Condition, ref="Light")
summary(f1 <- lm(f0raw ~ Condition, dat))

par(mfrow=c(2, 2))
hist(resid(f1))
qqnorm(resid(f1));qqline(resid(f1))
plot(fitted(f1), resid(f1))
```


## Free Speech

### Cycle Durations

```{r}
dat <- brm %>%
  filter(Role=="Confederate", Task=="Free")
```

```{r}
ggplot(dat, aes(Condition, cycleDur))+
  geom_boxplot()
```

* Cycles are shorter during biking than during sitting. There's a trend of Heavy being shorter than Light, but it doesn't reach significance.

```{r}
dat$Condition <- relevel(dat$Condition, ref="Light")
summary(b1 <- lm(cycleDur ~ Condition, dat))

par(mfrow=c(2, 2))
hist(resid(b1))
qqnorm(resid(b1));qqline(resid(b1))
plot(fitted(b1), resid(b1))
```

### Inhalation Amplitude

```{r}
dat <- brm %>%
  filter(Role=="Confederate", Task=="Free")
```

```{r}
ggplot(dat, aes(Condition, inhalAmp))+
  geom_boxplot()
```

* Inhalation amplitude gets larger with physical effort.

```{r}
dat$Condition <- relevel(dat$Condition, ref="Light")
summary(d1 <- lm(inhalAmp ~ Condition, dat))

par(mfrow=c(2, 2))
hist(resid(d1))
qqnorm(resid(d1));qqline(resid(d1))
plot(fitted(d1), resid(d1))
```

### F0

```{r}
dat <- fsm %>%
  filter(Task == "Free", Speaker == "Confederate")
```

```{r}
ggplot(dat, aes(Condition, f0raw))+
  geom_boxplot()
```

* F0 increased with physical effort.

```{r}
dat$Condition <- relevel(dat$Condition, ref="Light")
summary(m1 <- lm(f0raw ~ Condition, dat))

par(mfrow=c(2, 2))
hist(resid(m1))
qqnorm(resid(m1));qqline(resid(m1))
plot(fitted(m1), resid(m1))
```

# Participants

# 1 - Breathing. Is there a difference in respiratory behavior when people talk alone or in an interaction?

## Solo read speech

### Cycle Duration

```{r}
dat <- brm %>%
  filter(Role=="Participant", Task%in%c("ReadAlone", "ReadBaseline"))
```

```{r}
# Individual effects
ggplot(dat %>% filter(Condition %in% c("Baseline", "Sitting")), aes(Condition, cycleDur))+
  geom_boxplot()+
  facet_wrap(~Speaker, scales="free_y")+
  ggtitle("Solo Read - Cycle Duration")
ggsave(paste0(folderInd, "SoloRead-CycleDur-BaselineInteraction.png"), width = 3000, height=3000, units="px")
```

Many participants seemed to have shorter cycles during Sitting than during Baseline. A few of them had clearly longer cycles during Sitting than during Baseline.

```{r}
ggplot(dat %>% filter(Condition %in% c("Baseline", "Sitting")), aes(Condition, cycleDur))+
  geom_boxplot()

ggplot(dat, aes(Cond2, cycleDur))+
  geom_boxplot()
```

* Cycle duration doesn't change between Baseline and Sitting or Baseline and Interaction during solo read speech.

```{r}
summary(b1 <- lmer(cycleDur ~ Condition + (1+Condition| Speaker), dat %>% filter(Condition %in% c("Baseline", "Sitting"))))

summary(b1 <- lmer(cycleDur ~ Cond2 + (1+Cond2| Speaker), dat))

par(mfrow=c(2, 2))
hist(resid(b1))
qqnorm(resid(b1));qqline(resid(b1))
plot(fitted(b1), resid(b1))
```

### Inhalation Amplitude

```{r}
dat <- brm %>%
  filter(Role=="Participant", Task%in%c("ReadAlone", "ReadBaseline"))
```

```{r}
ggplot(dat %>% filter(Condition %in% c("Baseline", "Sitting")), aes(Condition, inhalAmp))+
  geom_boxplot()+
  facet_wrap(~Speaker, scales="free_y")+
  ggtitle("Solo Read - Inhalation amplitude")
ggsave(paste0(folderInd, "SoloRead-InhalAmp-BaselineInteraction.png"), width = 3000, height=3000, units="px")
```

Some participants didn't change. Half of them seemed to have lower inhalation amplitudes during Sitting, and around 5, the opposite.

```{r}
ggplot(dat %>% filter(Condition %in% c("Baseline", "Sitting")), aes(Condition, inhalAmp))+
  geom_boxplot()

ggplot(dat, aes(Cond2, inhalAmp))+
  geom_boxplot()
```

* Inhalation amplitude doesn't change between Baseline and Sitting or Baseline and Interaction during solo read speech.

```{r}
summary(b1 <- lmer(inhalAmp ~ Condition + (1+Condition| Speaker), dat %>% filter(Condition %in% c("Baseline", "Sitting"))))

summary(b1 <- lmer(inhalAmp ~ Cond2 + (1+Cond2| Speaker), dat))

par(mfrow=c(2, 2))
hist(resid(b1))
qqnorm(resid(b1));qqline(resid(b1))
plot(fitted(b1), resid(b1))
```

## Solo spontaneous speech

### Cycle Duration

```{r}
dat <- brm %>%
  filter(Role=="Participant", Task=="Free", act=="speaking")
```

```{r}
ggplot(dat %>% filter(Condition %in% c("Baseline", "Sitting")), aes(Condition, cycleDur))+
  geom_boxplot()+
  facet_wrap(~Speaker, scales="free_y")+
  ggtitle("Spontaneous - Cycle Duration")
ggsave(paste0(folderInd, "Spontaneous-CycleDur-BaselineInteraction.png"), width = 3000, height=3000, units="px")
```

A few participants had longer cycles during Sitting (vs. Baseline), a few the opposite. Most didn't show much of a difference.

```{r}
ggplot(dat %>% filter(Condition %in% c("Baseline", "Sitting")), aes(Condition, cycleDur))+
  geom_boxplot()

ggplot(dat, aes(Cond2, cycleDur))+
  geom_boxplot()
```

* Cycle duration doesn't change between Baseline and Sitting or Baseline and Interaction during solo spontaneous speech.

```{r}
summary(b1 <- lmer(cycleDur ~ Condition + (1+Condition| Speaker), dat %>% filter(Condition %in% c("Baseline", "Sitting"))))

summary(b1a <- lmer(cycleDur ~ Cond2 + (1+Cond2| Speaker), dat))

par(mfrow=c(2, 2))
hist(resid(b1))
qqnorm(resid(b1));qqline(resid(b1))
plot(fitted(b1), resid(b1))
```

### Inhalation Amplitude

```{r}
dat <- brm %>%
  filter(Role=="Participant", Task=="Free", act=="speaking")
```

```{r}
ggplot(dat %>% filter(Condition %in% c("Baseline", "Sitting")), aes(Condition, inhalAmp))+
  geom_boxplot()+
  facet_wrap(~Speaker, scales="free_y")+
  ggtitle("Spontaneous - Inhalation amplitude")
ggsave(paste0(folderInd, "Spontaneous-InhalAmp-BaselineInteraction.png"), width = 3000, height=3000, units="px")
```

A few speakers had deeper inhalations during Sitting, others the opposite. In general, not a big difference across conditions.

```{r}
ggplot(dat %>% filter(Condition %in% c("Baseline", "Sitting")), aes(Condition, inhalAmp))+
  geom_boxplot()

ggplot(dat, aes(Cond2, inhalAmp))+
  geom_boxplot()
```

* Inhalation amplitude doesn't change between Baseline and Sitting or Baseline and Interaction during solo spontaneous speech.

```{r}
summary(b1 <- lmer(inhalAmp ~ Condition + (1+Condition| Speaker), dat %>% filter(Condition %in% c("Baseline", "Sitting"))))

summary(b1 <- lmer(inhalAmp ~ Cond2 + (1+Cond2| Speaker), dat))

par(mfrow=c(2, 2))
hist(resid(b1))
qqnorm(resid(b1));qqline(resid(b1))
plot(fitted(b1), resid(b1))
```

# 1 - Speech. Is there a difference in speech behavior (f0) when people talk alone or in an interaction?

## Solo read speech

```{r}
dat <- frb %>%
  filter(Role=="Participant", Task%in%c("ReadAlone", "ReadBaseline"))
```

```{r}
ggplot(dat %>% filter(Condition %in% c("Baseline", "Sitting")), aes(Condition, f0raw))+
  geom_boxplot()+
  facet_wrap(~Speaker, scales="free_y")+
  ggtitle("Solo Read - F0")
ggsave(paste0(folderInd, "SoloRead-F0-BaselineInteraction.png"), width = 3000, height=3000, units="px")
```

Some participants had higher f0 during Sitting, only a couple the opposite. The difference between the conditions wasn't big.

```{r}
ggplot(dat %>% filter(Condition %in% c("Baseline", "Sitting")), aes(Condition, f0raw))+
  geom_boxplot()

ggplot(dat, aes(Cond2, f0raw))+
  geom_boxplot()
```

* F0 was higher during interaction than during baseline. (This is considering baseline vs. sitting AND baseline vs. interaction)

```{r}
summary(f1 <- lmer(f0raw ~ Condition + (1 + Condition | Speaker), dat %>% filter(Condition %in% c("Baseline", "Sitting"))))
summary(f1a <- lmer(f0raw ~ Cond2 + (1 + Cond2 | Speaker), dat))

par(mfrow=c(2, 2))
hist(resid(f1))
qqnorm(resid(f1));qqline(resid(f1))
plot(fitted(f1), resid(f1))

par(mfrow=c(2, 2))
hist(resid(f1a))
qqnorm(resid(f1a));qqline(resid(f1a))
plot(fitted(f1a), resid(f1a))
```

```{r}
dat <- frb %>% 
  filter(Task %in% c("ReadAlone", "ReadBaseline"), Condition %in% c("Baseline", "Sitting")) %>% 
  group_by(Speaker, Condition) %>% 
  mutate(f0mean = mean(f0raw, na.rm=TRUE)) %>% 
  ungroup()

ggplot(dat, aes(Condition, f0mean))+
  geom_point()+
  geom_line()+
  facet_wrap(~Speaker)
```

## Solo spontaneous speech

```{r}
dat <- fsm %>%
  filter(Role=="Participant", Task=="Free")
```

```{r}
ggplot(dat %>% filter(Condition %in% c("Baseline", "Sitting")), aes(Condition, f0raw))+
  geom_boxplot()+
  facet_wrap(~Speaker, scales="free_y")+
  ggtitle("Spontaneous - F0")
ggsave(paste0(folderInd, "Spontaneous-f0-BaselineInteraction.png"), width = 3000, height=3000, units="px")
```

A few had higher f0 during Sitting. Only a couple the opposite. Mostly no difference.

```{r}
ggplot(dat %>% filter(Condition %in% c("Baseline", "Sitting")), aes(Condition, f0raw))+
  geom_boxplot()

ggplot(dat, aes(Cond2, f0raw))+
  geom_boxplot()
```

* F0 didn't change from Baseline to Interaction OR from Baseline to Sitting.

```{r}
summary(f1 <- lmer(f0raw ~ Condition + (1 + Condition | Speaker), dat %>% filter(Condition %in% c("Baseline", "Sitting"))))
summary(f1a <- lmer(f0raw ~ Cond2 + (1 + Cond2 | Speaker), dat))

par(mfrow=c(2, 2))
hist(resid(f1))
qqnorm(resid(f1));qqline(resid(f1))
plot(fitted(f1), resid(f1))

par(mfrow=c(2, 2))
hist(resid(f1a))
qqnorm(resid(f1a));qqline(resid(f1a))
plot(fitted(f1a), resid(f1a))
```

# 2 - Breathing without overlapping speech

# 2.1 Do speakers adapt their speech breathing to an interlocutor who moves with different physical effort?

## Solo read speech

### Cycle Duration

```{r}
dat <- brm %>%
  filter(Role=="Participant", Task%in%c("ReadAlone", "ReadBaseline"), Condition!="Baseline")
```

```{r}
ggplot(dat, aes(Condition, cycleDur))+
  geom_boxplot()+
  facet_wrap(~Speaker, scales="free_y")+
  ggtitle("Solo Read - Cycle Duration")
ggsave(paste0(folderInd, "SoloRead-CycleDur-Conditions.png"), width = 3000, height=3000, units="px")
```

For some of the participants, Light seemed to be the most different condition. But in general, no difference across conditions for most participants.

```{r}
ggplot(dat, aes(Condition, cycleDur))+
  geom_boxplot()
```

* There is no effect of condition on cycle duration during solo read speech.

```{r}
dat$Condition <- relevel(dat$Condition, ref="Light")
summary(b1 <- lmer(cycleDur ~ Condition + (1+Condition| Speaker), dat))

par(mfrow=c(2, 2))
hist(resid(b1))
qqnorm(resid(b1));qqline(resid(b1))
plot(fitted(b1), resid(b1))
```

### Inhalation Amplitude

```{r}
dat <- brm %>%
  filter(Role=="Participant", Task%in%c("ReadAlone", "ReadBaseline"), Condition!="Baseline")
```

```{r}
ggplot(dat, aes(Condition, inhalAmp))+
  geom_boxplot()+
  facet_wrap(~Speaker, scales="free_y")+
  ggtitle("Solo Read - Inhalation amplitude")
ggsave(paste0(folderInd, "SoloRead-InhalAmp-Conditions.png"), width = 3000, height=3000, units="px")
```

For quite a few participants, inhalation amplitude seemed to go up with physical effort. For one, the opposite (but very markedly). For many, no difference.

```{r}
ggplot(dat, aes(Condition, inhalAmp))+
  geom_boxplot()
```

* There is no effect of condition on inhalation amplitude during solo read speech.

```{r}
dat$Condition <- relevel(dat$Condition, ref="Light")
summary(b1 <- lmer(inhalAmp ~ Condition + (1+Condition| Speaker), dat))

par(mfrow=c(2, 2))
hist(resid(b1))
qqnorm(resid(b1));qqline(resid(b1))
plot(fitted(b1), resid(b1))
```

## Solo spontaneous speech

### Cycle Duration

```{r}
dat <- brm %>%
  filter(Role=="Participant", Task=="Free", act=="speaking", Condition!="Baseline")
```

```{r}
ggplot(dat, aes(Condition, cycleDur))+
  geom_boxplot()+
  facet_wrap(~Speaker, scales="free_y")+
  ggtitle("Spontaneous - Cycle Duration")
ggsave(paste0(folderInd, "Spontaneous-CycleDur-Conditions.png"), width = 3000, height=3000, units="px")
```

For almost all participants, no difference.

```{r}
ggplot(dat, aes(Condition, cycleDur))+
  geom_boxplot()
```

* There is no effect of condition on cycle duration during solo spontaneous speech.

```{r}
dat$Condition <- relevel(dat$Condition, ref="Light")
summary(b1 <- lmer(cycleDur ~ Condition + (1+Condition| Speaker), dat))

par(mfrow=c(2, 2))
hist(resid(b1))
qqnorm(resid(b1));qqline(resid(b1))
plot(fitted(b1), resid(b1))
```

### Inhalation Amplitude

```{r}
dat <- brm %>%
  filter(Role=="Participant", Task=="Free", act=="speaking", Condition!="Baseline")
```

```{r}
ggplot(dat, aes(Condition, inhalAmp))+
  geom_boxplot()+
  facet_wrap(~Speaker, scales="free_y")+
  ggtitle("Spontaneous - Inhalation amplitude")
ggsave(paste0(folderInd, "Spontaneous-InhalAmp-Conditions.png"), width = 3000, height=3000, units="px")
```

For a few participants, there was either a decrease or increase in inhalation amplitude with physical effort. For a couple, it seemed non-linear. For the others, no difference.

```{r}
ggplot(dat, aes(Condition, inhalAmp))+
  geom_boxplot()
```

* There is no effect of condition on inhalation amplitude during solo spontaneous speech.

```{r}
dat$Condition <- relevel(dat$Condition, ref="Light")
summary(b1 <- lmer(inhalAmp ~ Condition + (1+Condition| Speaker), dat))

par(mfrow=c(2, 2))
hist(resid(b1))
qqnorm(resid(b1));qqline(resid(b1))
plot(fitted(b1), resid(b1))
```

# 2.2  Do speakers adapt their breathing while listening to an interlocutor who moves and talks with different physical effort?

### Cycle Duration

```{r}
dat <- brm %>%
  filter(Role=="Participant", act == "listening")
```

```{r}
ggplot(dat, aes(Condition, cycleDur))+
  geom_boxplot()+
  facet_wrap(~Speaker, scales="free_y")+
  ggtitle("Listening - Cycle Duration")
ggsave(paste0(folderInd, "Listening-CycleDur-Conditions.png"), width = 3000, height=3000, units="px")
```

No difference.

```{r}
ggplot(dat %>% filter(cycleDur < 7), aes(Condition, cycleDur))+
  geom_boxplot()
```

* Cycle duration doesn't change across conditions during listening.

```{r}
summary(c1 <- lmer(cycleDur ~ Condition + (1 + Condition |Speaker), dat))

par(mfrow=c(2, 2))
hist(resid(c1))
qqnorm(resid(c1));qqline(resid(c1))
plot(fitted(c1), resid(c1))
```

### Inhalation Amplitude

```{r}
dat <- brm %>%
  filter(Role=="Participant", act == "listening")
```

```{r}
ggplot(dat, aes(Condition, inhalAmp))+
  geom_boxplot()+
  facet_wrap(~Speaker, scales="free_y")+
  ggtitle("Listening - Inhalation amplitude")
ggsave(paste0(folderInd, "Listening-InhalAmp-Conditions.png"), width = 3000, height=3000, units="px")
```

For a couple of participants there was a clear in- or decrease in inhalation amplitude with physical effort. For some, non-linear effect (markedly). For the others, no difference.

```{r}
ggplot(dat %>% filter(cycleDur < 7), aes(Condition, inhalAmp))+
  geom_boxplot()
```

* Inhalation amplitude doesn't change across conditions during listening.

```{r}
summary(c1 <- lmer(inhalAmp ~ Condition + (1 + Condition |Speaker), dat))

par(mfrow=c(2, 2))
hist(resid(c1))
qqnorm(resid(c1));qqline(resid(c1))
plot(fitted(c1), resid(c1))
```

# 2 - Speech without overlap

# 2.1 Do speakers adapt their speech (f0) to an interlocutor who moves with different physical effort?

## Solo read speech

```{r}
dat <- frb %>%
  filter(Role=="Participant", Task=="ReadAlone")
```

```{r}
ggplot(dat, aes(Condition, f0raw))+
  geom_boxplot()+
  facet_wrap(~Speaker, scales="free_y")+
  ggtitle("Solo Read - F0")
ggsave(paste0(folderInd, "SoloRead-F0-Conditions.png"), width = 3000, height=3000, units="px")
```

For a couple of participants, increase with condition. In general, no difference.

```{r}
ggplot(dat, aes(Condition, f0raw))+
  geom_boxplot()
```

* F0 didn't change across conditions during solo read speech.

```{r}
summary(f1 <- lmer(f0raw ~ Condition + (1 + Condition | Speaker), dat))

par(mfrow=c(2, 2))
hist(resid(f1))
qqnorm(resid(f1));qqline(resid(f1))
plot(fitted(f1), resid(f1))
```

## Solo spontaneous speech

```{r}
dat <- fsm %>%
  filter(Role=="Participant", Task=="Free", Condition!="Baseline")
```

```{r}
ggplot(dat, aes(Condition, f0raw))+
  geom_boxplot()+
  facet_wrap(~Speaker, scales="free_y")+
  ggtitle("Spontaneous - F0")
ggsave(paste0(folderInd, "Spontaneous-F0-Conditions.png"), width = 3000, height=3000, units="px")
```

No difference.

```{r}
ggplot(dat, aes(Condition, f0raw))+
  geom_boxplot()
```

* F0 didn't change across conditions during solo free speech.

```{r}
summary(f1 <- lmer(f0raw ~ Condition + (1 + Condition | Speaker), dat))

par(mfrow=c(2, 2))
hist(resid(f1))
qqnorm(resid(f1));qqline(resid(f1))
plot(fitted(f1), resid(f1))
```

# 3 - Breathing during synchronous speech

# 3.1  Do speakers adapt their breathing while reading synchronously with an interlocutor in comparison to an interaction with no overlap?

### Cycle Duration

```{r}
dat <- brm %>%
  filter(Role=="Participant", Task %in% c("ReadJoint", "ReadAlone"))
```

```{r}
ggplot(dat, aes(Task, cycleDur))+
  geom_boxplot()+
  facet_wrap(~Speaker, scales="free_y")+
  ggtitle("Read Solo vs Sync - Cycle Duration")
ggsave(paste0(folderInd, "SoloSync-CycleDur-Task.png"), width = 3000, height=3000, units="px")
```

For a few, shorter cycles during sync read speech. In general, no difference.

```{r}
ggplot(dat, aes(Condition, cycleDur))+
  geom_boxplot()+
  facet_wrap(~Task)
```

* There is no effect of task (solo vs. synchronous) on cycle duration. And no interaction with condition.

```{r}
dat$Condition <- relevel(dat$Condition, ref="Light")

summary(b1 <- lmer(cycleDur ~ Task + (1+Task | Speaker), dat))

summary(b1a <- lmer(cycleDur ~ Task:Condition + (1+Task+Condition | Speaker), dat))

par(mfrow=c(2, 2))
hist(resid(b1))
qqnorm(resid(b1));qqline(resid(b1))
plot(fitted(b1), resid(b1))
```

### Inhalation Amplitude

```{r}
dat <- brm %>%
  filter(Role=="Participant", Task %in% c("ReadJoint", "ReadAlone"))
```

```{r}
ggplot(dat, aes(Task, inhalAmp))+
  geom_boxplot()+
  facet_wrap(~Speaker, scales="free_y")+
  ggtitle("Read Solo vs Sync - Inhalation amplitude")
ggsave(paste0(folderInd, "SoloSync-InhalAmp-Task.png"), width = 3000, height=3000, units="px")
```

No difference.

```{r}
ggplot(dat, aes(Condition, inhalAmp))+
  geom_boxplot()+
  facet_wrap(~Task)
```
 
* Inhalation amplitude was lower while reading synchronously than while reading solo. See model below for the interaction with Condition.

```{r}
dat$Condition <- relevel(dat$Condition, ref="Light")

summary(b1 <- lmer(inhalAmp ~ Task + (1+Task | Speaker), dat))

summary(b1a <- lmer(inhalAmp ~ Task:Condition + (1+Task +Condition| Speaker), dat))

par(mfrow=c(2, 2))
hist(resid(b1))
qqnorm(resid(b1));qqline(resid(b1))
plot(fitted(b1), resid(b1))
```

# 3.2 Do speakers adapt their breathing while reading synchronously with an interlocutor who moves and talks under different physical effort?

### Cycle Duration

```{r}
dat <- brm %>%
  filter(Role=="Participant", Task=="ReadJoint")
```

```{r}
ggplot(dat, aes(Condition, cycleDur))+
  geom_boxplot()+
  facet_wrap(~Speaker, scales="free_y")+
  ggtitle("Sync Read - Cycle duration")
ggsave(paste0(folderInd, "SyncRead-CycleDur-Conditions.png"), width = 3000, height=3000, units="px")
```

Slight change in cycle duration across condition for a couple of participants. In general, no difference.

```{r}
ggplot(dat, aes(Condition, cycleDur))+
  geom_boxplot()
```

* Cycle durations didn't change across conditions during synchronous read speech.

```{r}
dat$Condition <- relevel(dat$Condition, ref="Light")

summary(b1 <- lmer(cycleDur ~ Condition + (1 +Condition| Speaker), dat))

par(mfrow=c(2, 2))
hist(resid(b1))
qqnorm(resid(b1));qqline(resid(b1))
plot(fitted(b1), resid(b1))
```

### Inhalation Amplitude

```{r}
dat <- brm %>%
  filter(Role=="Participant", Task=="ReadJoint")
```

```{r}
ggplot(dat, aes(Condition, inhalAmp))+
  geom_boxplot()+
  facet_wrap(~Speaker, scales="free_y")+
  ggtitle("Sync Read - Inhalation amplitude")
ggsave(paste0(folderInd, "SyncRead-InhalAmp-Conditions.png"), width = 3000, height=3000, units="px")
```

For a few, in- or decrease with condition; others, a clear non-linear effect. For many, no difference.

```{r}
ggplot(dat, aes(Condition, inhalAmp))+
  geom_boxplot()
```

* Inhalation amplitude was larger during Light Biking than during Sitting and Heavy Biking. No difference between Sitting and Heavy Biking.

```{r}
dat$Condition <- relevel(dat$Condition, ref="Light")

summary(b1 <- lmer(inhalAmp ~ Condition + (1 | Speaker), dat))

par(mfrow=c(2, 2))
hist(resid(b1))
qqnorm(resid(b1));qqline(resid(b1))
plot(fitted(b1), resid(b1))
```

# 3 - Synchronous speech

# 3.1  Do speakers adapt their speech (f0) while reading synchronously with an interlocutor in comparison to an interaction with no overlap?

```{r}
dat <- frb %>%
  filter(Role=="Participant", Task %in% c("ReadJoint", "ReadAlone"))
```

```{r}
ggplot(dat, aes(Task, f0raw))+
  geom_boxplot()+
  facet_wrap(~Speaker, scales="free_y")+
  ggtitle("Read Solo vs. Sync - F0")
ggsave(paste0(folderInd, "SyncRead-F0-Task.png"), width = 3000, height=3000, units="px")
```

For a few, higher f0 during sync read speech; for a couple, the opposite. In general, no difference.

```{r}
ggplot(dat, aes(Condition, f0raw))+
  geom_boxplot()+
  facet_wrap(~Task)
```

* There is no effect of task (solo vs. synchronous) on f0. See model below for the interaction with condition.

```{r}
dat$Condition <- relevel(dat$Condition, ref="Light")

summary(b1 <- lmer(f0raw ~ Task + (1+Task | Speaker), dat))

summary(b1 <- lmer(f0raw ~ Task:Condition + (1+Task+Condition | Speaker), dat))

par(mfrow=c(2, 2))
hist(resid(b1))
qqnorm(resid(b1));qqline(resid(b1))
plot(fitted(b1), resid(b1))
```

# 3.2 Do speakers adapt their speech (f0) while reading synchronously with an interlocutor who moves and talks under different physical effort?

```{r}
dat <- frb %>%
  filter(Role=="Participant", Task=="ReadJoint")
```

```{r}
ggplot(dat, aes(Condition, f0raw))+
  geom_boxplot()+
  facet_wrap(~Speaker, scales="free_y")+
  ggtitle("Sync Read - F0")
ggsave(paste0(folderInd, "SyncRead-F0-Conditions.png"), width = 3000, height=3000, units="px")
```

For most, increase with condition.

```{r}
ggplot(dat, aes(Condition, f0raw))+
  geom_boxplot()
```

* F0 is higher during Heavy Biking than Sitting or Light Biking.
* There is no significant difference between Light and Sitting, but a tendency for Light to have higher f0 (following the pattern).

```{r}
dat$Condition <- relevel(dat$Condition, ref="Light")

summary(b1 <- lmer(f0raw ~ Condition + (1 + Condition | Speaker), dat))

par(mfrow=c(2, 2))
hist(resid(b1))
qqnorm(resid(b1));qqline(resid(b1))
plot(fitted(b1), resid(b1))
```