---
title: "Analysis - f0 median (to compare with previous results)"
author: "Tom Offrede"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(lme4)
library(tidyverse)
library(broom.mixed)

folder <- "C:/Users/offredet/Documents/1HU/ExperimentBreathing/Data/DataForAnalysis/AllData/"
folderInd <- "C:/Users/offredet/Documents/1HU/ExperimentBreathing/Data/IndividualEffects/" # to save plots with individual effects

orderCond <- c("Sitting", "Light", "Heavy")
orderCondBase <- c("Baseline", orderCond)
orderAct <- c("watching", "listening")

load(paste0(folder, "DataReadSpeech-Median.RData"))
```

# Confederate

## Read Speech

### F0

```{r}
dat <- frbMed %>%
  filter(grepl("Read", Task), Speaker == "Confederate")
```

```{r}
ggplot(dat, aes(Condition, f0raw))+
  geom_boxplot()
```

# Participants

# Do speakers adapt their speech (f0) to an interlocutor who moves with different physical effort?

## Solo read speech

```{r}
dat <- frbMed %>%
  filter(Role=="Participant", Task=="ReadAlone")
```

```{r}
ggplot(dat, aes(Condition, f0raw))+
  geom_boxplot()+
  facet_wrap(~Speaker, scales="free_y")+
  ggtitle("Solo Read - F0")
# ggsave(paste0(folderInd, "SoloRead-F0-Conditions.png"), width = 3000, height=3000, units="px")
```

For a couple of participants, increase with condition. In general, no difference.

```{r}
ggplot(dat, aes(Condition, f0raw))+
  geom_boxplot()
```

* F0 didn't change across conditions during solo read speech.

```{r}
summary(f1 <- lmer(f0raw ~ Condition + (1 + Condition | Speaker), dat))

par(mfrow=c(2, 2))
hist(resid(f1))
qqnorm(resid(f1));qqline(resid(f1))
plot(fitted(f1), resid(f1))
```

## Synchronous read speech

```{r}
dat <- frbMed %>%
  filter(Role=="Participant", Task=="ReadJoint")
```

```{r}
ggplot(dat, aes(Condition, f0raw))+
  geom_boxplot()+
  facet_wrap(~Speaker, scales="free_y")+
  ggtitle("Sync Read - F0")
ggsave(paste0(folderInd, "SyncRead-F0-Conditions.png"), width = 3000, height=3000, units="px")
```

For most, increase with condition.

```{r}
ggplot(dat, aes(Condition, f0raw))+
  geom_boxplot()
```

* F0 is higher during Heavy Biking than Sitting or Light Biking.
* There is no significant difference between Light and Sitting, but a tendency for Light to have higher f0 (following the pattern).

```{r}
# dat$Condition <- relevel(dat$Condition, ref="Light")

summary(b1 <- lmer(f0raw ~ Condition + (1 + Condition | Speaker), dat))

par(mfrow=c(2, 2))
hist(resid(b1))
qqnorm(resid(b1));qqline(resid(b1))
plot(fitted(b1), resid(b1))
```

