---
title: "Analysis - SMC Paper Structured"
author: "Tom Offrede"
date: "2023-01-27"
output:
  html_document:
    toc: true
    toc_float: 
      collapsed: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, include=FALSE}
library(lme4)
library(tidyverse)
library(broom.mixed)

folder <- "C:/Users/offredet/Documents/1HU/ExperimentBreathing/Data/DataForAnalysis/AllData/"
folderInd <- "C:/Users/offredet/Documents/1HU/ExperimentBreathing/Data/IndividualEffects/" # to save plots with individual effects

orderCond <- c("Sitting", "Light", "Heavy")
orderCondBase <- c("Baseline", orderCond)
orderAct <- c("watching", "listening")

load(paste0(folder, "DataBreathing.RData"))
load(paste0(folder, "DataReadSpeech.RData"))
load(paste0(folder, "DataSpeech.RData"))
load(paste0(folder, "DataConfBaselineBreathing.RData"))
```

# Raw values for appendix

## Cycle Duration

### Confederate

```{r}
# Read Speech Solo
## Sitting
mean(brm$cycleDur[brm$Speaker=="Confederate" & grepl("Read", brm$Task) & brm$Condition=="Sitting"], na.rm=TRUE) # 2.72
median(brm$cycleDur[brm$Speaker=="Confederate" & grepl("Read", brm$Task) & brm$Condition=="Sitting"], na.rm=TRUE) # 2.48
sd(brm$cycleDur[brm$Speaker=="Confederate" & grepl("Read", brm$Task) & brm$Condition=="Sitting"], na.rm=TRUE) # 1.11

## Light
mean(brm$cycleDur[brm$Speaker=="Confederate" & grepl("Read", brm$Task) & brm$Condition=="Light"], na.rm=TRUE) # 3.06
median(brm$cycleDur[brm$Speaker=="Confederate" & grepl("Read", brm$Task) & brm$Condition=="Light"], na.rm=TRUE) # 3.03
sd(brm$cycleDur[brm$Speaker=="Confederate" & grepl("Read", brm$Task) & brm$Condition=="Light"], na.rm=TRUE) # 0.85

## Heavy
mean(brm$cycleDur[brm$Speaker=="Confederate" & grepl("Read", brm$Task) & brm$Condition=="Heavy"], na.rm=TRUE) # 2.7
median(brm$cycleDur[brm$Speaker=="Confederate" & grepl("Read", brm$Task) & brm$Condition=="Heavy"], na.rm=TRUE) # 2.78
sd(brm$cycleDur[brm$Speaker=="Confederate" & grepl("Read", brm$Task) & brm$Condition=="Heavy"], na.rm=TRUE) # 0.68

# Watching (silent)
## Sitting
mean(bc$cycleDur[bc$condition=="Sitting"], na.rm=TRUE) # 2.98
median(bc$cycleDur[bc$condition=="Sitting"], na.rm=TRUE) # 2.88
sd(bc$cycleDur[bc$condition=="Sitting"], na.rm=TRUE) # 0.75

## Light
mean(bc$cycleDur[bc$condition=="Light"], na.rm=TRUE) # 2.57
median(bc$cycleDur[bc$condition=="Light"], na.rm=TRUE) # 2.43
sd(bc$cycleDur[bc$condition=="Light"], na.rm=TRUE) # 0.77

## Heavy
mean(bc$cycleDur[bc$condition=="Heavy"], na.rm=TRUE) # 2.25
median(bc$cycleDur[bc$condition=="Heavy"], na.rm=TRUE) # 2.23
sd(bc$cycleDur[bc$condition=="Heavy"], na.rm=TRUE) # 0.61
```

### Participants

```{r}
# Read Speech Solo
## Baseline
mean(brm$cycleDur[brm$Speaker!="Confederate" & brm$Task=="ReadBaseline" & brm$Condition=="Baseline"], na.rm=TRUE) # 4.77
median(brm$cycleDur[brm$Speaker!="Confederate" & brm$Task=="ReadBaseline" & brm$Condition=="Baseline"], na.rm=TRUE) # 4.32
sd(brm$cycleDur[brm$Speaker!="Confederate" & brm$Task=="ReadBaseline" & brm$Condition=="Baseline"], na.rm=TRUE) # 2.46

## Sitting
mean(brm$cycleDur[brm$Speaker!="Confederate" & brm$Task=="ReadAlone" & brm$Condition=="Sitting"], na.rm=TRUE) # 4.59
median(brm$cycleDur[brm$Speaker!="Confederate" & brm$Task=="ReadAlone" & brm$Condition=="Sitting"], na.rm=TRUE) # 4.2
sd(brm$cycleDur[brm$Speaker!="Confederate" & brm$Task=="ReadAlone" & brm$Condition=="Sitting"], na.rm=TRUE) # 2.23

## Light
mean(brm$cycleDur[brm$Speaker!="Confederate" & brm$Task=="ReadAlone" & brm$Condition=="Light"], na.rm=TRUE) # 4.66
median(brm$cycleDur[brm$Speaker!="Confederate" & brm$Task=="ReadAlone" & brm$Condition=="Light"], na.rm=TRUE) # 4.24
sd(brm$cycleDur[brm$Speaker!="Confederate" & brm$Task=="ReadAlone" & brm$Condition=="Light"], na.rm=TRUE) # 2.35

## Heavy
mean(brm$cycleDur[brm$Speaker!="Confederate" & brm$Task=="ReadAlone" & brm$Condition=="Heavy"], na.rm=TRUE) # 4.59
median(brm$cycleDur[brm$Speaker!="Confederate" & brm$Task=="ReadAlone" & brm$Condition=="Heavy"], na.rm=TRUE) # 4.27
sd(brm$cycleDur[brm$Speaker!="Confederate" & brm$Task=="ReadAlone" & brm$Condition=="Heavy"], na.rm=TRUE) # 2.1

# Read Speech Sync
## Sitting
mean(brm$cycleDur[brm$Speaker!="Confederate" & brm$Task=="ReadJoint" & brm$Condition=="Sitting"], na.rm=TRUE) # 4.58
median(brm$cycleDur[brm$Speaker!="Confederate" & brm$Task=="ReadJoint" & brm$Condition=="Sitting"], na.rm=TRUE) # 4.08
sd(brm$cycleDur[brm$Speaker!="Confederate" & brm$Task=="ReadJoint" & brm$Condition=="Sitting"], na.rm=TRUE) # 2.19

## Light
mean(brm$cycleDur[brm$Speaker!="Confederate" & brm$Task=="ReadJoint" & brm$Condition=="Light"], na.rm=TRUE) # 4.54
median(brm$cycleDur[brm$Speaker!="Confederate" & brm$Task=="ReadJoint" & brm$Condition=="Light"], na.rm=TRUE) # 3.89
sd(brm$cycleDur[brm$Speaker!="Confederate" & brm$Task=="ReadJoint" & brm$Condition=="Light"], na.rm=TRUE) # 2.51

## Heavy
mean(brm$cycleDur[brm$Speaker!="Confederate" & brm$Task=="ReadJoint" & brm$Condition=="Heavy"], na.rm=TRUE) # 4.45
median(brm$cycleDur[brm$Speaker!="Confederate" & brm$Task=="ReadJoint" & brm$Condition=="Heavy"], na.rm=TRUE) # 3.86
sd(brm$cycleDur[brm$Speaker!="Confederate" & brm$Task=="ReadJoint" & brm$Condition=="Heavy"], na.rm=TRUE) # 2.27

# Watching (silent)
## Sitting
mean(brm$cycleDur[brm$Speaker!="Confederate" & brm$act=="watching" & brm$Condition=="Sitting"], na.rm=TRUE) # 3.85
median(brm$cycleDur[brm$Speaker!="Confederate" & brm$act=="watching" & brm$Condition=="Sitting"], na.rm=TRUE) # 3.43
sd(brm$cycleDur[brm$Speaker!="Confederate" & brm$act=="watching" & brm$Condition=="Sitting"], na.rm=TRUE) # 1.7

## Light
mean(brm$cycleDur[brm$Speaker!="Confederate" & brm$act=="watching" & brm$Condition=="Light"], na.rm=TRUE) # 4.03
median(brm$cycleDur[brm$Speaker!="Confederate" & brm$act=="watching" & brm$Condition=="Light"], na.rm=TRUE) # 3.42
sd(brm$cycleDur[brm$Speaker!="Confederate" & brm$act=="watching" & brm$Condition=="Light"], na.rm=TRUE) # 2.4

## Heavy
mean(brm$cycleDur[brm$Speaker!="Confederate" & brm$act=="watching" & brm$Condition=="Heavy"], na.rm=TRUE) # 3.89
median(brm$cycleDur[brm$Speaker!="Confederate" & brm$act=="watching" & brm$Condition=="Heavy"], na.rm=TRUE) # 3.31
sd(brm$cycleDur[brm$Speaker!="Confederate" & brm$act=="watching" & brm$Condition=="Heavy"], na.rm=TRUE) # 1.96
```

## Inahlation Amplitude

### Confederate

```{r}
# Read Speech Solo
## Sitting
mean(brm$inhalAmp[brm$Speaker=="Confederate" & grepl("Read", brm$Task) & brm$Condition=="Sitting"], na.rm=TRUE) # 0.28
median(brm$inhalAmp[brm$Speaker=="Confederate" & grepl("Read", brm$Task) & brm$Condition=="Sitting"], na.rm=TRUE) # 0.27
sd(brm$inhalAmp[brm$Speaker=="Confederate" & grepl("Read", brm$Task) & brm$Condition=="Sitting"], na.rm=TRUE) # 0.12

## Light
mean(brm$inhalAmp[brm$Speaker=="Confederate" & grepl("Read", brm$Task) & brm$Condition=="Light"], na.rm=TRUE) # 0.51
median(brm$inhalAmp[brm$Speaker=="Confederate" & grepl("Read", brm$Task) & brm$Condition=="Light"], na.rm=TRUE) # 0.49
sd(brm$inhalAmp[brm$Speaker=="Confederate" & grepl("Read", brm$Task) & brm$Condition=="Light"], na.rm=TRUE) # 0.18

## Heavy
mean(brm$inhalAmp[brm$Speaker=="Confederate" & grepl("Read", brm$Task) & brm$Condition=="Heavy"], na.rm=TRUE) # 0.6
median(brm$inhalAmp[brm$Speaker=="Confederate" & grepl("Read", brm$Task) & brm$Condition=="Heavy"], na.rm=TRUE) # 0.63
sd(brm$inhalAmp[brm$Speaker=="Confederate" & grepl("Read", brm$Task) & brm$Condition=="Heavy"], na.rm=TRUE) # 0.16

# Watching (silent)
## Sitting
mean(bc$inhalAmp[bc$condition=="Sitting"], na.rm=TRUE) # 0.38
median(bc$inhalAmp[bc$condition=="Sitting"], na.rm=TRUE) # 0.34
sd(bc$inhalAmp[bc$condition=="Sitting"], na.rm=TRUE) # 0.16

## Light
mean(bc$inhalAmp[bc$condition=="Light"], na.rm=TRUE) # 0.37
median(bc$inhalAmp[bc$condition=="Light"], na.rm=TRUE) # 0.32
sd(bc$inhalAmp[bc$condition=="Light"], na.rm=TRUE) # 0.17

## Heavy
mean(bc$inhalAmp[bc$condition=="Heavy"], na.rm=TRUE) # 0.37
median(bc$inhalAmp[bc$condition=="Heavy"], na.rm=TRUE) # 0.37
sd(bc$inhalAmp[bc$condition=="Heavy"], na.rm=TRUE) # 0.15
```

### Participants

```{r}
# Read Speech Solo
## Baseline
mean(brm$inhalAmp[brm$Speaker!="Confederate" & brm$Task=="ReadBaseline" & brm$Condition=="Baseline"], na.rm=TRUE) # 0.42
median(brm$inhalAmp[brm$Speaker!="Confederate" & brm$Task=="ReadBaseline" & brm$Condition=="Baseline"], na.rm=TRUE) # 0.39
sd(brm$inhalAmp[brm$Speaker!="Confederate" & brm$Task=="ReadBaseline" & brm$Condition=="Baseline"], na.rm=TRUE) # 0.19

## Sitting
mean(brm$inhalAmp[brm$Speaker!="Confederate" & brm$Task=="ReadAlone" & brm$Condition=="Sitting"], na.rm=TRUE) # 0.38
median(brm$inhalAmp[brm$Speaker!="Confederate" & brm$Task=="ReadAlone" & brm$Condition=="Sitting"], na.rm=TRUE) # 0.36
sd(brm$inhalAmp[brm$Speaker!="Confederate" & brm$Task=="ReadAlone" & brm$Condition=="Sitting"], na.rm=TRUE) # 0.21

## Light
mean(brm$inhalAmp[brm$Speaker!="Confederate" & brm$Task=="ReadAlone" & brm$Condition=="Light"], na.rm=TRUE) # 0.4
median(brm$inhalAmp[brm$Speaker!="Confederate" & brm$Task=="ReadAlone" & brm$Condition=="Light"], na.rm=TRUE) # 0.36
sd(brm$inhalAmp[brm$Speaker!="Confederate" & brm$Task=="ReadAlone" & brm$Condition=="Light"], na.rm=TRUE) # 0.19

## Heavy
mean(brm$inhalAmp[brm$Speaker!="Confederate" & brm$Task=="ReadAlone" & brm$Condition=="Heavy"], na.rm=TRUE) # 0.41
median(brm$inhalAmp[brm$Speaker!="Confederate" & brm$Task=="ReadAlone" & brm$Condition=="Heavy"], na.rm=TRUE) # 0.38
sd(brm$inhalAmp[brm$Speaker!="Confederate" & brm$Task=="ReadAlone" & brm$Condition=="Heavy"], na.rm=TRUE) # 0.2

# Read Speech Sync
## Sitting
mean(brm$inhalAmp[brm$Speaker!="Confederate" & brm$Task=="ReadJoint" & brm$Condition=="Sitting"], na.rm=TRUE) # 0.36
median(brm$inhalAmp[brm$Speaker!="Confederate" & brm$Task=="ReadJoint" & brm$Condition=="Sitting"], na.rm=TRUE) # 0.35
sd(brm$inhalAmp[brm$Speaker!="Confederate" & brm$Task=="ReadJoint" & brm$Condition=="Sitting"], na.rm=TRUE) # 0.18

## Light
mean(brm$inhalAmp[brm$Speaker!="Confederate" & brm$Task=="ReadJoint" & brm$Condition=="Light"], na.rm=TRUE) # 0.39
median(brm$inhalAmp[brm$Speaker!="Confederate" & brm$Task=="ReadJoint" & brm$Condition=="Light"], na.rm=TRUE) # 0.36
sd(brm$inhalAmp[brm$Speaker!="Confederate" & brm$Task=="ReadJoint" & brm$Condition=="Light"], na.rm=TRUE) # 0.2

## Heavy
mean(brm$inhalAmp[brm$Speaker!="Confederate" & brm$Task=="ReadJoint" & brm$Condition=="Heavy"], na.rm=TRUE) # 0.36
median(brm$inhalAmp[brm$Speaker!="Confederate" & brm$Task=="ReadJoint" & brm$Condition=="Heavy"], na.rm=TRUE) # 0.32
sd(brm$inhalAmp[brm$Speaker!="Confederate" & brm$Task=="ReadJoint" & brm$Condition=="Heavy"], na.rm=TRUE) # 0.2

# Watching (silent)
## Sitting
mean(brm$inhalAmp[brm$Speaker!="Confederate" & brm$act=="watching" & brm$Condition=="Sitting"], na.rm=TRUE) # 0.47
median(brm$inhalAmp[brm$Speaker!="Confederate" & brm$act=="watching" & brm$Condition=="Sitting"], na.rm=TRUE) # 0.45
sd(brm$inhalAmp[brm$Speaker!="Confederate" & brm$act=="watching" & brm$Condition=="Sitting"], na.rm=TRUE) # 0.22

## Light
mean(brm$inhalAmp[brm$Speaker!="Confederate" & brm$act=="watching" & brm$Condition=="Light"], na.rm=TRUE) # 0.4
median(brm$inhalAmp[brm$Speaker!="Confederate" & brm$act=="watching" & brm$Condition=="Light"], na.rm=TRUE) # 0.39
sd(brm$inhalAmp[brm$Speaker!="Confederate" & brm$act=="watching" & brm$Condition=="Light"], na.rm=TRUE) # 0.21

## Heavy
mean(brm$inhalAmp[brm$Speaker!="Confederate" & brm$act=="watching" & brm$Condition=="Heavy"], na.rm=TRUE) # 0.39
median(brm$inhalAmp[brm$Speaker!="Confederate" & brm$act=="watching" & brm$Condition=="Heavy"], na.rm=TRUE) # 0.36
sd(brm$inhalAmp[brm$Speaker!="Confederate" & brm$act=="watching" & brm$Condition=="Heavy"], na.rm=TRUE) # 0.21
```

## f0

### Confederate

```{r}
# Read Speech Solo
## Sitting
mean(frb$f0raw[frb$Speaker=="Confederate" & grepl("Read", frb$Task) & frb$Condition=="Sitting"], na.rm=TRUE) # 200.96
median(frb$f0raw[frb$Speaker=="Confederate" & grepl("Read", frb$Task) & frb$Condition=="Sitting"], na.rm=TRUE) # 199.6
sd(frb$f0raw[frb$Speaker=="Confederate" & grepl("Read", frb$Task) & frb$Condition=="Sitting"], na.rm=TRUE) # 16.1

## Light
mean(frb$f0raw[frb$Speaker=="Confederate" & grepl("Read", frb$Task) & frb$Condition=="Light"], na.rm=TRUE) # 209.3
median(frb$f0raw[frb$Speaker=="Confederate" & grepl("Read", frb$Task) & frb$Condition=="Light"], na.rm=TRUE) # 210.1
sd(frb$f0raw[frb$Speaker=="Confederate" & grepl("Read", frb$Task) & frb$Condition=="Light"], na.rm=TRUE) # 14.72

## Heavy
mean(frb$f0raw[frb$Speaker=="Confederate" & grepl("Read", frb$Task) & frb$Condition=="Heavy"], na.rm=TRUE) # 220.41
median(frb$f0raw[frb$Speaker=="Confederate" & grepl("Read", frb$Task) & frb$Condition=="Heavy"], na.rm=TRUE) # 221.75
sd(frb$f0raw[frb$Speaker=="Confederate" & grepl("Read", frb$Task) & frb$Condition=="Heavy"], na.rm=TRUE) # 13.08
```

### Participants

```{r}
# Read Speech Solo
## Baseline
mean(frb$f0raw[frb$Speaker!="Confederate" & frb$Task=="ReadBaseline" & frb$Condition=="Baseline"], na.rm=TRUE) # 200.56
median(frb$f0raw[frb$Speaker!="Confederate" & frb$Task=="ReadBaseline" & frb$Condition=="Baseline"], na.rm=TRUE) # 204.87
sd(frb$f0raw[frb$Speaker!="Confederate" & frb$Task=="ReadBaseline" & frb$Condition=="Baseline"], na.rm=TRUE) # 30.1

## Sitting
mean(frb$f0raw[frb$Speaker!="Confederate" & frb$Task=="ReadAlone" & frb$Condition=="Sitting"], na.rm=TRUE) # 208.2
median(frb$f0raw[frb$Speaker!="Confederate" & frb$Task=="ReadAlone" & frb$Condition=="Sitting"], na.rm=TRUE) # 210.2
sd(frb$f0raw[frb$Speaker!="Confederate" & frb$Task=="ReadAlone" & frb$Condition=="Sitting"], na.rm=TRUE) # 23.86

## Light
mean(frb$f0raw[frb$Speaker!="Confederate" & frb$Task=="ReadAlone" & frb$Condition=="Light"], na.rm=TRUE) # 208.63
median(frb$f0raw[frb$Speaker!="Confederate" & frb$Task=="ReadAlone" & frb$Condition=="Light"], na.rm=TRUE) # 211.08
sd(frb$f0raw[frb$Speaker!="Confederate" & frb$Task=="ReadAlone" & frb$Condition=="Light"], na.rm=TRUE) # 27.26

## Heavy
mean(frb$f0raw[frb$Speaker!="Confederate" & frb$Task=="ReadAlone" & frb$Condition=="Heavy"], na.rm=TRUE) # 209.49
median(frb$f0raw[frb$Speaker!="Confederate" & frb$Task=="ReadAlone" & frb$Condition=="Heavy"], na.rm=TRUE) # 210.7
sd(frb$f0raw[frb$Speaker!="Confederate" & frb$Task=="ReadAlone" & frb$Condition=="Heavy"], na.rm=TRUE) # 24.56

# Read Speech Sync
## Sitting
mean(frb$f0raw[frb$Speaker!="Confederate" & frb$Task=="ReadJoint" & frb$Condition=="Sitting"], na.rm=TRUE) # 206.21
median(frb$f0raw[frb$Speaker!="Confederate" & frb$Task=="ReadJoint" & frb$Condition=="Sitting"], na.rm=TRUE) # 209.74
sd(frb$f0raw[frb$Speaker!="Confederate" & frb$Task=="ReadJoint" & frb$Condition=="Sitting"], na.rm=TRUE) # 28.24

## Light
mean(frb$f0raw[frb$Speaker!="Confederate" & frb$Task=="ReadJoint" & frb$Condition=="Light"], na.rm=TRUE) # 208.13
median(frb$f0raw[frb$Speaker!="Confederate" & frb$Task=="ReadJoint" & frb$Condition=="Light"], na.rm=TRUE) # 211.58
sd(frb$f0raw[frb$Speaker!="Confederate" & frb$Task=="ReadJoint" & frb$Condition=="Light"], na.rm=TRUE) # 27.42

## Heavy
mean(frb$f0raw[frb$Speaker!="Confederate" & frb$Task=="ReadJoint" & frb$Condition=="Heavy"], na.rm=TRUE) # 211.57
median(frb$f0raw[frb$Speaker!="Confederate" & frb$Task=="ReadJoint" & frb$Condition=="Heavy"], na.rm=TRUE) # 214.12
sd(frb$f0raw[frb$Speaker!="Confederate" & frb$Task=="ReadJoint" & frb$Condition=="Heavy"], na.rm=TRUE) # 27.02
```


# Confederate

## Read Speech

### Cycle Durations

```{r}
dat <- brm %>%
  filter(Role=="Confederate", grepl("Read", Task))
```

```{r}
ggplot(dat, aes(Condition, cycleDur))+
  geom_boxplot()
```

* Difference only between Light and the other conditions, but no difference between Sitting and Heavy. So there is probably noise in the data and we can consider that there was no effect of condition on breathing cycle duration.

```{r}
# dat$Condition <- relevel(dat$Condition, ref="Light")
summary(b1 <- lm(cycleDur ~ Condition, dat))

par(mfrow=c(2, 2))
hist(resid(b1))
qqnorm(resid(b1));qqline(resid(b1))
plot(fitted(b1), resid(b1))
```

### Inhalation Amplitude

```{r}
dat <- brm %>%
  filter(Role=="Confederate", grepl("Read", Task))
```

```{r}
ggplot(dat, aes(Condition, inhalAmp))+
  geom_boxplot()
```

* Inhalation amplitude got higher with physical effort.

```{r}
# dat$Condition <- relevel(dat$Condition, ref="Light")
summary(d1 <- lm(inhalAmp ~ Condition, dat))

par(mfrow=c(2, 2))
hist(resid(d1))
qqnorm(resid(d1));qqline(resid(d1))
plot(fitted(d1), resid(d1))
```

### F0

```{r}
dat <- frb %>%
  filter(grepl("Read", Task), Speaker == "Confederate")
```

```{r}
ggplot(dat, aes(Condition, f0raw))+
  geom_boxplot()
```

* F0 got higher with physical effort.

```{r}
# dat$Condition <- relevel(dat$Condition, ref="Light")
summary(f1 <- lm(f0raw ~ Condition, dat))

par(mfrow=c(2, 2))
hist(resid(f1))
qqnorm(resid(f1));qqline(resid(f1))
plot(fitted(f1), resid(f1))
```

# Participants

# Breathing. Do people adapt their breathing to a confederate who is moving under different levels of physical effort when both of them are silent?

## Cycle Duration

```{r}
dat <- brm %>% 
  filter(Role=="Participant", act=="watching")
```

```{r}
# Individual effects
ggplot(dat, aes(Condition, cycleDur))+
  geom_boxplot()+
  facet_wrap(~Speaker, scales="free_y")+
  ggtitle("Watching - Cycle Duration")
ggsave(paste0(folderInd, "Watching-CycleDur.png"), width = 3000, height=3000, units="px")
```

A few participants behaved in opposite ways.

```{r}
ggplot(dat, aes(Condition, cycleDur))+
  geom_boxplot()
```

No effect of condition on cycle duration while watching.

```{r}
summary(b1 <- lmer(cycleDur ~ Condition + (1+Condition| Speaker), dat))

par(mfrow=c(2, 2))
hist(resid(b1))
qqnorm(resid(b1));qqline(resid(b1))
plot(fitted(b1), resid(b1))
```


## Inhalation Amplitude

```{r}
dat <- brm %>% 
  filter(Role=="Participant", act=="watching")
```

```{r}
# Individual effects
ggplot(dat, aes(Condition, inhalAmp))+
  geom_boxplot()+
  facet_wrap(~Speaker, scales="free_y")+
  ggtitle("Watching - Inhalation Amplitude")
ggsave(paste0(folderInd, "Watching-InhalAmp.png"), width = 3000, height=3000, units="px")
```

Again, a few participants behaved in opposite ways.

```{r}
ggplot(dat, aes(Condition, inhalAmp))+
  geom_boxplot()
```

Lower inhalation amplitude during Heavy than Sitting, with Light in between.

```{r}
summary(b1 <- lmer(inhalAmp ~ Condition + (1+Condition| Speaker), dat))

par(mfrow=c(2, 2))
hist(resid(b1))
qqnorm(resid(b1));qqline(resid(b1))
plot(fitted(b1), resid(b1))
```

# Do speakers adapt their speech breathing to an interlocutor who moves with different physical effort?

## Solo read speech

### Cycle Duration

```{r}
dat <- brm %>%
  filter(Role=="Participant", Task%in%c("ReadAlone", "ReadBaseline"), Condition!="Baseline")
```

```{r}
ggplot(dat, aes(Condition, cycleDur))+
  geom_boxplot()+
  facet_wrap(~Speaker, scales="free_y")+
  ggtitle("Solo Read - Cycle Duration")
ggsave(paste0(folderInd, "SoloRead-CycleDur-Conditions.png"), width = 3000, height=3000, units="px")
```

For some of the participants, Light seemed to be the most different condition. But in general, no difference across conditions for most participants.

```{r}
ggplot(dat, aes(Condition, cycleDur))+
  geom_boxplot()
```

* There is no effect of condition on cycle duration during solo read speech.

```{r}
# dat$Condition <- relevel(dat$Condition, ref="Light")
summary(b1 <- lmer(cycleDur ~ Condition + (1+Condition| Speaker), dat))

par(mfrow=c(2, 2))
hist(resid(b1))
qqnorm(resid(b1));qqline(resid(b1))
plot(fitted(b1), resid(b1))
```

### Inhalation Amplitude

```{r}
dat <- brm %>%
  filter(Role=="Participant", Task%in%c("ReadAlone", "ReadBaseline"), Condition!="Baseline")
```

```{r}
ggplot(dat, aes(Condition, inhalAmp))+
  geom_boxplot()+
  facet_wrap(~Speaker, scales="free_y")+
  ggtitle("Solo Read - Inhalation amplitude")
ggsave(paste0(folderInd, "SoloRead-InhalAmp-Conditions.png"), width = 3000, height=3000, units="px")
```

For quite a few participants, inhalation amplitude seemed to go up with physical effort. For one, the opposite (but very markedly). For many, no difference.

```{r}
ggplot(dat, aes(Condition, inhalAmp))+
  geom_boxplot()
```

* There is no effect of condition on inhalation amplitude during solo read speech.

```{r}
# dat$Condition <- relevel(dat$Condition, ref="Light")
summary(b1 <- lmer(inhalAmp ~ Condition + (1+Condition| Speaker), dat))

par(mfrow=c(2, 2))
hist(resid(b1))
qqnorm(resid(b1));qqline(resid(b1))
plot(fitted(b1), resid(b1))
```



## Synchronous read speech

### Cycle Duration

```{r}
dat <- brm %>%
  filter(Role=="Participant", Task=="ReadJoint")
```

```{r}
ggplot(dat, aes(Condition, cycleDur))+
  geom_boxplot()+
  facet_wrap(~Speaker, scales="free_y")+
  ggtitle("Sync Read - Cycle duration")
ggsave(paste0(folderInd, "SyncRead-CycleDur-Conditions.png"), width = 3000, height=3000, units="px")
```

Slight change in cycle duration across condition for a couple of participants. In general, no difference.

```{r}
ggplot(dat, aes(Condition, cycleDur))+
  geom_boxplot()
```

* Cycle durations didn't change across conditions during synchronous read speech.

```{r}
# dat$Condition <- relevel(dat$Condition, ref="Light")

summary(b1 <- lmer(cycleDur ~ Condition + (1 +Condition| Speaker), dat))

par(mfrow=c(2, 2))
hist(resid(b1))
qqnorm(resid(b1));qqline(resid(b1))
plot(fitted(b1), resid(b1))
```

### Inhalation Amplitude

```{r}
dat <- brm %>%
  filter(Role=="Participant", Task=="ReadJoint")
```

```{r}
ggplot(dat, aes(Condition, inhalAmp))+
  geom_boxplot()+
  facet_wrap(~Speaker, scales="free_y")+
  ggtitle("Sync Read - Inhalation amplitude")
ggsave(paste0(folderInd, "SyncRead-InhalAmp-Conditions.png"), width = 3000, height=3000, units="px")
```

For a few, in- or decrease with condition; others, a clear non-linear effect. For many, no difference.

```{r}
ggplot(dat, aes(Condition, inhalAmp))+
  geom_boxplot()
```

* Inhalation amplitude was larger during Light Biking than during Sitting and Heavy Biking. No difference between Sitting and Heavy Biking.

```{r}
# dat$Condition <- relevel(dat$Condition, ref="Light")

summary(b1 <- lmer(inhalAmp ~ Condition + (1 | Speaker), dat))

par(mfrow=c(2, 2))
hist(resid(b1))
qqnorm(resid(b1));qqline(resid(b1))
plot(fitted(b1), resid(b1))
```


# Do speakers adapt their speech (f0) to an interlocutor who moves with different physical effort?

## Solo read speech

```{r}
dat <- frb %>%
  filter(Role=="Participant", Task=="ReadAlone")
```

```{r}
ggplot(dat, aes(Condition, f0raw))+
  geom_boxplot()+
  facet_wrap(~Speaker, scales="free_y")+
  ggtitle("Solo Read - F0")
ggsave(paste0(folderInd, "SoloRead-F0-Conditions.png"), width = 3000, height=3000, units="px")
```

For a couple of participants, increase with condition. In general, no difference.

```{r}
ggplot(dat, aes(Condition, f0raw))+
  geom_boxplot()
```

* F0 didn't change across conditions during solo read speech.

```{r}
summary(f1 <- lmer(f0raw ~ Condition + (1 + Condition | Speaker), dat))

par(mfrow=c(2, 2))
hist(resid(f1))
qqnorm(resid(f1));qqline(resid(f1))
plot(fitted(f1), resid(f1))
```

## Synchronous read speech

```{r}
dat <- frb %>%
  filter(Role=="Participant", Task=="ReadJoint")
```

```{r}
ggplot(dat, aes(Condition, f0raw))+
  geom_boxplot()+
  facet_wrap(~Speaker, scales="free_y")+
  ggtitle("Sync Read - F0")
ggsave(paste0(folderInd, "SyncRead-F0-Conditions.png"), width = 3000, height=3000, units="px")
```

For most, increase with condition.

```{r}
ggplot(dat, aes(Condition, f0raw))+
  geom_boxplot()
```

* F0 is higher during Heavy Biking than Sitting or Light Biking.
* There is no significant difference between Light and Sitting, but a tendency for Light to have higher f0 (following the pattern).

```{r}
# dat$Condition <- relevel(dat$Condition, ref="Light")

summary(b1 <- lmer(f0raw ~ Condition + (1 + Condition | Speaker), dat))

par(mfrow=c(2, 2))
hist(resid(b1))
qqnorm(resid(b1));qqline(resid(b1))
plot(fitted(b1), resid(b1))
```

# Individual Differences

```{r}
# ind <- data.frame(matrix(ncol=6, nrow=0))
# names(ind) <- c("speaker", "task", "variable", "SL", "LH", "SH")
# 
# dat <- brm %>% 
#   filter(Role=="Participant", Task=="ReadAlone")
# dat1 <- frb %>% 
#   filter(Role=="Participant", Task=="ReadAlone")
# 
# for(s in unique(dat$Speaker)){
#   dat$Condition <- relevel(dat$Condition, ref="Sitting")
#   c <- tidy(lm(cycleDur ~ Condition, dat %>% filter(Speaker==s)))
#   ifelse(c$statistic[c$term=="ConditionLight"] >= 2, SL <- "pos", ifelse(c$statistic[c$term=="ConditionLight"] <= -2, SL <- "neg", SL <- NA))
#   ifelse(c$statistic[c$term=="ConditionHeavy"] >= 2, SH <- "pos", ifelse(c$statistic[c$term=="ConditionHeavy"] <= -2, SH <- "neg", SH <- NA))
#   dat$Condition <- relevel(dat$Condition, ref="Light")
#   c <- tidy(lm(cycleDur ~ Condition, dat %>% filter(Speaker==s)))
#   ifelse(c$statistic[c$term=="ConditionHeavy"] >= 2, LH <- "pos", ifelse(c$statistic[c$term=="ConditionHeavy"] <= -2, LH <- "neg", LH <- NA))
#   ind[nrow(ind)+1,] <- c(s,
#                          "Solo",
#                          "cycleDur",
#                          SL, LH, SH)
#   
#   dat$Condition <- relevel(dat$Condition, ref="Sitting")
#   c <- tidy(lm(inhalAmp ~ Condition, dat %>% filter(Speaker==s)))
#   ifelse(c$statistic[c$term=="ConditionLight"] >= 2, SL <- "pos", ifelse(c$statistic[c$term=="ConditionLight"] <= -2, SL <- "neg", SL <- NA))
#   ifelse(c$statistic[c$term=="ConditionHeavy"] >= 2, SH <- "pos", ifelse(c$statistic[c$term=="ConditionHeavy"] <= -2, SH <- "neg", SH <- NA))
#   dat$Condition <- relevel(dat$Condition, ref="Light")
#   c <- tidy(lm(inhalAmp ~ Condition, dat %>% filter(Speaker==s)))
#   ifelse(c$statistic[c$term=="ConditionHeavy"] >= 2, LH <- "pos", ifelse(c$statistic[c$term=="ConditionHeavy"] <= -2, LH <- "neg", LH <- NA))
#   ind[nrow(ind)+1,] <- c(s,
#                          "Solo",
#                          "inhalAmp",
#                          SL, LH, SH)
#   
#   dat1$Condition <- relevel(dat1$Condition, ref="Sitting")
#   c <- tidy(lm(f0raw ~ Condition, dat1 %>% filter(Speaker==s)))
#   ifelse(c$statistic[c$term=="ConditionLight"] >= 2, SL <- "pos", ifelse(c$statistic[c$term=="ConditionLight"] <= -2, SL <- "neg", SL <- NA))
#   ifelse(c$statistic[c$term=="ConditionHeavy"] >= 2, SH <- "pos", ifelse(c$statistic[c$term=="ConditionHeavy"] <= -2, SH <- "neg", SH <- NA))
#   dat1$Condition <- relevel(dat1$Condition, ref="Light")
#   c <- tidy(lm(f0raw ~ Condition, dat1 %>% filter(Speaker==s)))
#   ifelse(c$statistic[c$term=="ConditionHeavy"] >= 2, LH <- "pos", ifelse(c$statistic[c$term=="ConditionHeavy"] <= -2, LH <- "neg", LH <- NA))
#   ind[nrow(ind)+1,] <- c(s,
#                          "Solo",
#                          "f0",
#                          SL, LH, SH)
# }
# 
# dat <- brm %>% 
#   filter(Role=="Participant", Task=="ReadJoint")
# dat1 <- frb %>% 
#   filter(Role=="Participant", Task=="ReadJoint")
# 
# summary(lm(inhalAmp ~ Condition, dat %>% filter(Speaker=="TSE")))
# 
# for(s in unique(dat$Speaker)){
#   dat$Condition <- relevel(dat$Condition, ref="Sitting")
#   c <- tidy(lm(cycleDur ~ Condition, dat %>% filter(Speaker==s)))
#   ifelse(c$statistic[c$term=="ConditionLight"] >= 2, SL <- "pos", ifelse(c$statistic[c$term=="ConditionLight"] <= -2, SL <- "neg", SL <- NA))
#   ifelse(c$statistic[c$term=="ConditionHeavy"] >= 2, SH <- "pos", ifelse(c$statistic[c$term=="ConditionHeavy"] <= -2, SH <- "neg", SH <- NA))
#   dat$Condition <- relevel(dat$Condition, ref="Light")
#   c <- tidy(lm(cycleDur ~ Condition, dat %>% filter(Speaker==s)))
#   ifelse(c$statistic[c$term=="ConditionHeavy"] >= 2, LH <- "pos", ifelse(c$statistic[c$term=="ConditionHeavy"] <= -2, LH <- "neg", LH <- NA))
#   ind[nrow(ind)+1,] <- c(s,
#                          "Sync",
#                          "cycleDur",
#                          SL, LH, SH)
#   
#   dat$Condition <- relevel(dat$Condition, ref="Sitting")
#   c <- tidy(lm(inhalAmp ~ Condition, dat %>% filter(Speaker==s)))
#   ifelse(c$statistic[c$term=="ConditionLight"] >= 2, SL <- "pos", ifelse(c$statistic[c$term=="ConditionLight"] <= -2, SL <- "neg", SL <- NA))
#   ifelse(c$statistic[c$term=="ConditionHeavy"] >= 2, SH <- "pos", ifelse(c$statistic[c$term=="ConditionHeavy"] <= -2, SH <- "neg", SH <- NA))
#   dat$Condition <- relevel(dat$Condition, ref="Light")
#   c <- tidy(lm(inhalAmp ~ Condition, dat %>% filter(Speaker==s)))
#   ifelse(c$statistic[c$term=="ConditionHeavy"] >= 2, LH <- "pos", ifelse(c$statistic[c$term=="ConditionHeavy"] <= -2, LH <- "neg", LH <- NA))
#   ind[nrow(ind)+1,] <- c(s,
#                          "Sync",
#                          "inhalAmp",
#                          SL, LH, SH)
#   
#   dat1$Condition <- relevel(dat1$Condition, ref="Sitting")
#   c <- tidy(lm(f0raw ~ Condition, dat1 %>% filter(Speaker==s)))
#   ifelse(c$statistic[c$term=="ConditionLight"] >= 2, SL <- "pos", ifelse(c$statistic[c$term=="ConditionLight"] <= -2, SL <- "neg", SL <- NA))
#   ifelse(c$statistic[c$term=="ConditionHeavy"] >= 2, SH <- "pos", ifelse(c$statistic[c$term=="ConditionHeavy"] <= -2, SH <- "neg", SH <- NA))
#   dat1$Condition <- relevel(dat1$Condition, ref="Light")
#   c <- tidy(lm(f0raw ~ Condition, dat1 %>% filter(Speaker==s)))
#   ifelse(c$statistic[c$term=="ConditionHeavy"] >= 2, LH <- "pos", ifelse(c$statistic[c$term=="ConditionHeavy"] <= -2, LH <- "neg", LH <- NA))
#   ind[nrow(ind)+1,] <- c(s,
#                          "Sync",
#                          "f0",
#                          SL, LH, SH)
# }
# 
# dat <- brm %>% 
#   filter(Role=="Participant", act=="watching")
# 
# for(s in unique(dat$Speaker)){
#   dat$Condition <- relevel(dat$Condition, ref="Sitting")
#   c <- tidy(lm(cycleDur ~ Condition, dat %>% filter(Speaker==s)))
#   ifelse(c$statistic[c$term=="ConditionLight"] >= 2, SL <- "pos", ifelse(c$statistic[c$term=="ConditionLight"] <= -2, SL <- "neg", SL <- NA))
#   ifelse(c$statistic[c$term=="ConditionHeavy"] >= 2, SH <- "pos", ifelse(c$statistic[c$term=="ConditionHeavy"] <= -2, SH <- "neg", SH <- NA))
#   dat$Condition <- relevel(dat$Condition, ref="Light")
#   c <- tidy(lm(cycleDur ~ Condition, dat %>% filter(Speaker==s)))
#   ifelse(c$statistic[c$term=="ConditionHeavy"] >= 2, LH <- "pos", ifelse(c$statistic[c$term=="ConditionHeavy"] <= -2, LH <- "neg", LH <- NA))
#   ind[nrow(ind)+1,] <- c(s,
#                          "Watch",
#                          "cycleDur",
#                          SL, LH, SH)
#   
#   dat$Condition <- relevel(dat$Condition, ref="Sitting")
#   c <- tidy(lm(inhalAmp ~ Condition, dat %>% filter(Speaker==s)))
#   ifelse(c$statistic[c$term=="ConditionLight"] >= 2, SL <- "pos", ifelse(c$statistic[c$term=="ConditionLight"] <= -2, SL <- "neg", SL <- NA))
#   ifelse(c$statistic[c$term=="ConditionHeavy"] >= 2, SH <- "pos", ifelse(c$statistic[c$term=="ConditionHeavy"] <= -2, SH <- "neg", SH <- NA))
#   dat$Condition <- relevel(dat$Condition, ref="Light")
#   c <- tidy(lm(inhalAmp ~ Condition, dat %>% filter(Speaker==s)))
#   ifelse(c$statistic[c$term=="ConditionHeavy"] >= 2, LH <- "pos", ifelse(c$statistic[c$term=="ConditionHeavy"] <= -2, LH <- "neg", LH <- NA))
#   ind[nrow(ind)+1,] <- c(s,
#                          "Watch",
#                          "inhalAmp",
#                          SL, LH, SH)
# }
# 
# ind <- ind %>% 
#   filter_at(vars(SL,SH,LH),any_vars(!is.na(.))) # keep rows with at least one non-NA value
```

```{r}
# WD <- ind %>% 
#   filter(task=="Watch", variable=="cycleDur")
# 
# WI <- ind %>% 
#   filter(task=="Watch", variable=="inhalAmp")
# 
# SoI <- ind%>% 
#   filter(task=="Solo", variable=="inhalAmp")
# 
# SoD <- ind%>% 
#   filter(task=="Solo", variable=="cycleDur")
# 
# SoF <- ind%>% 
#   filter(task=="Solo", variable=="f0")
# 
# SyI <- ind%>% 
#   filter(task=="Sync", variable=="inhalAmp")
# 
# SyD <- ind%>% 
#   filter(task=="Sync", variable=="cycleDur")
# 
# SyF <- ind%>% 
#   filter(task=="Sync", variable=="f0")
```

