---
title: "Analysis-Defense"
author: "Tom Offrede"
date: "2024-04-29"
output:
  html_document:
    toc: true
    toc_float: 
      collapsed: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
Sys.setenv(LANG="en")
```

```{r, include=FALSE}
library(lmerTest)
library(tidyverse)
library(broom.mixed)
`%!in%` <- Negate(`%in%`)

folder <- "C:/Users/offredet/Documents/1HU/ExperimentBreathing/Data/DataForAnalysis/AllData/"
folderInd <- "C:/Users/offredet/Documents/1HU/ExperimentBreathing/Defense/" # to save plots with individual effects

orderCond <- c("Sitting", "Light", "Heavy")
orderCondBase <- c("Baseline", orderCond)
orderAct <- c("watching", "listening")

load(paste0(folder, "DataBreathing2.RData"))
load(paste0(folder, "DataConfBaselineBreathing.RData"))

brm <- brm2 |> rename(speaker = Speaker, condition = Condition) |> 
  mutate(Ifraction = inhalDur / cycleDur,
         act2 = ifelse(grepl("Read", act), "read", ifelse(grepl("speak", act), "speak", act))) |> # I-fraction is defined by Wilder & Baker 1974 as "the ratio of the length of inspiration to the total duration of the respiratory cycle"
  filter(speaker!="Confederate")
brm <- droplevels(brm)
brc0 <- brm2 |> 
  rename(speaker = Speaker, condition = Condition) |> 
  filter(speaker=="Confederate") |> 
  select(names(bc))
brc <- rbind(brc0, bc) |> 
  mutate(Ifraction = inhalDur / cycleDur)
brc <- droplevels(brc)
remove(brm2)
```

# Watching vs Listening

## Cycle duration

Breath cycles are shorter during listening, i.e. faster respiration rate. This is in line with previous literature (listening breathing is somewhere between quiet and speech breathing).

```{r}
ggplot(brm |> filter(act%in%c("watching", "listening"), cycleDur<7.5), aes(act, cycleDur))+
  geom_boxplot()+
  scale_x_discrete(limits=orderAct)

summary(lmer(cycleDur ~ act + (1+act|speaker), brm |> filter(act%in%c("watching", "listening"))))
```

**No condition effect during listening**

```{r}
ggplot(brm |> filter(act%in%c("listening"), cycleDur<7.5), aes(condition, cycleDur))+
  geom_boxplot()

summary(lmer(cycleDur ~ condition + (1|speaker), brm |> filter(act%in%c("listening"))))
```

## Inhalation amplitude

Inhalation amplitude is much lower while listening than watching

```{r}
ggplot(brm |> filter(#!is.na(act)
  act%in%c("watching", "listening")
  ), aes(act, inhalAmp))+
  geom_boxplot()+
  scale_x_discrete(limits=orderAct)

summary(lmer(inhalAmp ~ act + (1+act|speaker), brm |> filter(act%in%c("watching", "listening"))))
```

**No condition effect during listening**

```{r}
ggplot(brm |> filter(act%in%c("listening")), aes(condition, inhalAmp))+
  geom_boxplot()

summary(lmer(inhalAmp ~ condition + (1+condition|speaker), brm |> filter(act%in%c("listening"))))
```

## I-fraction

```{r}
ggplot(brm |> filter(#!is.na(act)
  act%in%c("watching", "listening")
  ), aes(act, Ifraction))+
  geom_boxplot()

summary(lmer(Ifraction ~ act + (1+act|speaker), brm |> filter(act%in%c("watching", "listening"))))
```

**No condition effect for listening nor watching**

```{r}
ggplot(brm |> filter(#!is.na(act)
  act%in%c("watching", "listening")
  ), aes(condition, Ifraction))+
  geom_boxplot()+
  facet_wrap(~act)

summary(lmer(Ifraction ~ condition + (1|speaker), brm |> filter(act%in%c("listening"))))

summary(lmer(Ifraction ~ condition + (1|speaker), brm |> filter(act%in%c("watching"))))
```

# Confederate

## Cycle duration

In read speech, we had found a weird pattern for cycle duration (reported in paper).

In spontaneous speech, the expected pattern is found: faster breathing with higher effort (although, no difference between Light and Heavy).

Overall, breathing also tends to be faster during spontaneous than read speech, but only slightly so (p = 0.042).

```{r}
brc$condition <- relevel(brc$condition, ref="Light")

ggplot(brc|> filter(act!="watching"), aes(condition, cycleDur))+
  geom_boxplot()+
  facet_wrap(~act)+
  scale_x_discrete(limits=orderCond)

summary(lm(cycleDur ~ condition, brc |> filter(act=="ReadAlone")))
summary(lm(cycleDur ~ condition, brc |> filter(act=="speaking")))

ggplot(brc|> filter(act!="watching"), aes(act, cycleDur))+
  geom_boxplot()

summary(lm(cycleDur ~ act, brc |> filter(act!="watching")))
```

## Inhalation amplitude

Inhalation amplitude increased more with condition during read speech than with spontaneous speech (during sitting there was no difference between read and spontaneous). In other words: overall, amplitude is larger in read than spontaneous speech; and this is accounted for by the larger effect of condition on amplitude in read speech.
 
```{r}
ggplot(brc|> filter(act!="watching"), aes(condition, inhalAmp))+
  geom_boxplot()+
  facet_wrap(~act)+
  scale_x_discrete(limits=orderCond)

summary(lm(inhalAmp ~ condition, brc |> filter(act=="ReadAlone")))
summary(lm(inhalAmp ~ condition, brc |> filter(act=="speaking")))

summary(m <- lm(inhalAmp ~ condition * act, brc |> filter(act%in%c("speaking", "ReadAlone"))))
summary(emmeans::emmeans(m, pairwise ~ act|condition))

ggplot(brc|> filter(act!="watching"), aes(act, inhalAmp))+
  geom_boxplot()

summary(lm(inhalAmp ~ act, brc |> filter(act!="watching")))
```


## I-fraction

**New thoughts about I-fraction: in this context, I'm not so interested in this variable, because what we could learn from it we could probably learn with more accuracy from inhalation duration**

In both read and free speech, I-fraction generally increases with increase in effort (although no difference between Light and Heavy during spontaneous speech, as was the case for cycle duration; and no difference between Sitting and Light during read speech).

While biking, I-fraction is higher in spontaneous speech.

(During sitting, I-fraction was slightly higher during read speech.)

Remember: higher I-fraction = more balanced inhalation / exhalation durations.
 
```{r}
ggplot(brc|> filter(act!="watching"), aes(condition, Ifraction))+
  geom_boxplot()+
  facet_wrap(~act)+
  scale_x_discrete(limits=orderCond)

m <- lm(Ifraction ~ condition * act, brc |> filter(act!="watching"))
summary(emmeans::emmeans(m, pairwise ~ condition|act))
summary(emmeans::emmeans(m, pairwise ~ act|condition))


ggplot(brc|> filter(act!="watching"), aes(act, Ifraction))+
  geom_boxplot()

summary(lm(Ifraction ~ act, brc |> filter(act!="watching")))
```


# Participants: exploring read vs spontaneous speech

## General comparisons

### Cycle duration

* Free speech: slower breathing rate during biking conditions, as compared to sitting (and sitting was same as baseline)

* Read speech: no condition effect in either alone nor joint. Also no overall difference between baseline, alone and joint.

* Free vs read: overall, longer cycles (slower rate) during free, than read, speech

```{r}
brm$condition <- relevel(brm$condition, ref="Light")

ggplot(brm|> filter(act%!in%c("watching", "listening", "ReadBaseline", "speakingBaseline")), aes(condition, cycleDur))+
  geom_boxplot()+
  facet_wrap(~act)+
  scale_x_discrete(limits=orderCond)

summary(lmer(cycleDur ~ condition + (1|speaker), brm |> filter(act=="ReadAlone")))
summary(lmer(cycleDur ~ condition + (1|speaker), brm |> filter(act=="ReadJoint")))
summary(lmer(cycleDur ~ condition + (1|speaker), brm |> filter(act=="speakingConditions")))

ggplot(brm|> filter(act%!in%c("watching", "listening")), aes(act, cycleDur))+
  geom_boxplot()

summary(lmer(cycleDur ~ act + (1|speaker), brm |> filter(grepl("Read|speak", act))))
```


### Inhalation amplitude

* Free speech: no robust condition differences.

* Read speech: **alone**: no condition difference. **joint**: light with higher amplitude (already in paper).

* Free vs read: higher amplitudes at baseline and alone, and lower at joint and free. (overall, also a decreasing pattern)

```{r}
brm$condition <- relevel(brm$condition, ref="Light")
brm$act <- relevel(brm$act, ref="ReadJoint")

ggplot(brm|> filter(act%!in%c("watching", "listening", "ReadBaseline", "speakingBaseline")), aes(condition, inhalAmp))+
  geom_boxplot()+
  facet_wrap(~act)+
  scale_x_discrete(limits=orderCond)

summary(lmer(inhalAmp ~ condition + (1+condition|speaker), brm |> filter(act=="ReadAlone")))
summary(lmer(inhalAmp ~ condition + (1|speaker), brm |> filter(act=="ReadJoint")))
summary(lmer(inhalAmp ~ condition + (1+condition|speaker), brm |> filter(act=="speakingConditions")))

ggplot(brm|> filter(act%!in%c("watching", "listening")), aes(act, inhalAmp))+
  geom_boxplot()+
  scale_x_discrete(limits=c("ReadBaseline", "ReadAlone", "ReadJoint", "speakingBaseline", "speakingConditions"))

summary(lmer(inhalAmp ~ act + (1+act|speaker), brm |> filter(grepl("Read|speak", act))))
```


### I-fraction

* Free speech: Sitting has larger I-fraction than Heavy (Light is not different from either)

* Read speech: **alone**: no condition difference. **joint**: higher I-fraction with higher effort (but no difference between light and heavy)

* Free vs read: no overall differences, just a few specific ones which may have been due to sample size (e.g. I-fraction larger in speakingConditions than speakingBaseline, and in ReadJoint than in all other categories)


```{r}
brm$condition <- relevel(brm$condition, ref="Sitting")

ggplot(brm|> filter(act%!in%c("watching", "listening", "ReadBaseline", "speakingBaseline")), aes(condition, Ifraction))+
  geom_boxplot()+
  facet_wrap(~act)+
  scale_x_discrete(limits=orderCond)

summary(lmer(Ifraction ~ condition + (1|speaker), brm |> filter(act=="ReadAlone")))
summary(lmer(Ifraction ~ condition + (1|speaker), brm |> filter(act=="ReadJoint")))
summary(lmer(Ifraction ~ condition + (1|speaker), brm |> filter(act=="speakingConditions")))

ggplot(brm|> filter(act%!in%c("watching", "listening")), aes(act, Ifraction))+
  geom_boxplot()+
  scale_x_discrete(limits=c("ReadBaseline", "ReadAlone", "ReadJoint", "speakingBaseline", "speakingConditions"))

brm$act <- relevel(brm$act, ref="ReadAlone")
summary(lmer(Ifraction ~ act + (1|speaker), brm |> filter(grepl("Read|speak", act))))

ggplot(brm|> filter(act2%in%c("read", "speak")), aes(act2, Ifraction))+
  geom_boxplot()

summary(lmer(Ifraction ~ act2 + (1+act2|speaker), brm |> filter(grepl("read|speak", act2))))
```

## Comparison: speakBaseline vs ReadBaseline

### Cycle duration

* Longer cycles (slower breathing rate) during free speech

```{r}
ggplot(brm |> filter(act%in%c("speakingBaseline", "ReadBaseline")), aes(act, cycleDur))+
  geom_boxplot()

summary(lmer(cycleDur ~ act + (1+act|speaker), brm |> filter(act%in%c("speakingBaseline", "ReadBaseline"))))
```

### Inhalation amplitude

* Lower inhalation amplitudes in free speech

```{r}
ggplot(brm |> filter(act%in%c("speakingBaseline", "ReadBaseline")), aes(act, inhalAmp))+
  geom_boxplot()

summary(lmer(inhalAmp ~ act + (1+act|speaker), brm |> filter(act%in%c("speakingBaseline", "ReadBaseline"))))
```

### I-fraction

* No difference in I-fraction

Interesting, given the robust differences in cycle duration and inhalation amplitude!

```{r}
ggplot(brm |> filter(act%in%c("speakingBaseline", "ReadBaseline")), aes(act, Ifraction))+
  geom_boxplot()

summary(lmer(Ifraction ~ act + (1+act|speaker), brm |> filter(act%in%c("speakingBaseline", "ReadBaseline"))))
```

## Comparison: speakConditions vs ReadAlone

### Cycle duration

* Longer cycles (slower breathing rate) during free speech (same as baseline)

```{r}
ggplot(brm |> filter(act%in%c("speakingConditions", "ReadAlone")), aes(#act, cycleDur))+
  geom_boxplot()

summary(lmer(cycleDur ~ act + (1+act|speaker), brm |> filter(act%in%c("speakingConditions", "ReadAlone"))))
```

* Read: no condition difference

* Free: sitting -- shorter cycle duration (faster rate) than biking conditions

```{r}
ggplot(brm |> filter(act%in%c("speakingConditions", "ReadAlone")), aes(condition, cycleDur))+
  geom_boxplot()+
  facet_wrap(~act)

summary(lmer(cycleDur ~ condition + (1|speaker), brm |> filter(act%in%c("ReadAlone"))))

brm$condition <- relevel(brm$condition, ref="Light")
summary(lmer(cycleDur ~ condition + (1|speaker), brm |> filter(act%in%c("speakingConditions"))))
```

### Inhalation amplitude

* Tendency to lower inhalation amplitudes in free speech (like baseline), but difference not significant

```{r}
ggplot(brm |> filter(act%in%c("speakingConditions", "ReadAlone")), aes(act, inhalAmp))+
  geom_boxplot()

summary(lmer(inhalAmp ~ act + (1+act|speaker), brm |> filter(act%in%c("speakingConditions", "ReadAlone"))))
```

* Read: no condition difference

* Free: no condition difference

```{r}
ggplot(brm |> filter(act%in%c("speakingConditions", "ReadAlone")), aes(condition, inhalAmp))+
  geom_boxplot()+
  facet_wrap(~act)

summary(lmer(inhalAmp ~ condition + (1|speaker), brm |> filter(act%in%c("ReadAlone"))))

brm$condition <- relevel(brm$condition, ref="Light")
summary(lmer(inhalAmp ~ condition + (1|speaker), brm |> filter(act%in%c("speakingConditions"))))
```

### I-fraction

* No difference in I-fraction (same as baseline)

```{r}
ggplot(brm |> filter(act%in%c("speakingConditions", "ReadAlone")), aes(act, Ifraction))+
  geom_boxplot()

summary(lmer(Ifraction ~ act + (1+act|speaker), brm |> filter(act%in%c("speakingConditions", "ReadAlone"))))
```

* Read: no condition difference

* Free: no condition difference

```{r}
ggplot(brm |> filter(act%in%c("speakingConditions", "ReadAlone")), aes(condition, Ifraction))+
  geom_boxplot()+
  facet_wrap(~act)

summary(lmer(Ifraction ~ condition + (1|speaker), brm |> filter(act%in%c("ReadAlone"))))

brm$condition <- relevel(brm$condition, ref="Light")
summary(lmer(Ifraction ~ condition + (1|speaker), brm |> filter(act%in%c("speakingConditions"))))
```

## Since (almost) no difference between conditions, conflate Baseline and Alone

```{r}
dat <- brm |> 
  mutate_at(c("act"), as.character) |> 
  mutate(act = ifelse(act %in% c("speakingBaseline", "speakingConditions"), "free", ifelse(act %in% c("ReadBaseline", "ReadAlone"), "read", NA))) |> 
  filter(!is.na(act)) |> 
  mutate_at(c("act"), as.factor)
```

### Cycle duration

* Longer cycles (slower breathing rate) during free speech

```{r}
ggplot(dat, aes(act, cycleDur))+
  geom_boxplot()

ggplot(dat, aes(act, cycleDur))+
  geom_boxplot()+
  facet_wrap(~speaker)

summary(lmer(cycleDur ~ act + (1+act|speaker), dat))
```

### Inhalation amplitude

* Lower inhalation amplitudes in free speech

```{r}
ggplot(dat, aes(act, inhalAmp))+
  geom_boxplot()

ggplot(dat, aes(act, inhalAmp))+
  geom_boxplot()+
  facet_wrap(~speaker)

summary(lmer(inhalAmp ~ act + (1+act|speaker), dat))
```

### I-fraction

### I-fraction

* No difference in I-fraction

Makes sense, given that both inhalation AND exhalation durations were longer in free speech. So participants divide their breath cycles equally regardless of the type of speech.

```{r}
ggplot(dat, aes(act, Ifraction))+
  geom_boxplot()

ggplot(dat, aes(act, Ifraction))+
  geom_boxplot()+
  facet_wrap(~speaker)

summary(lmer(Ifraction ~ act + (1+act|speaker), dat))
```

