---
title: "Breathing analysis"
output: html_document
date: '2022-05-05'
---

```{r, include=FALSE}
library(lme4)
library(tidyverse)
```


```{r setup, include=FALSE}
folder <- "C:/Users/tomof/Documents/1HU/ExperimentBreathing/Data/DataForAnalysis/AllData/"
load(paste0(folder, "BreathingData.RData"))

br$Condition[substr(br$file, 1, 1) == "B"] <- "Baseline"
br$Condition[substr(br$file, 1, 1) == "S"] <- "Sitting"
br$Condition[substr(br$file, 1, 1) == "L"] <- "Light"
br$Condition[substr(br$file, 1, 1) == "H"] <- "Heavy"
br$Speaker <- substr(br$file, 4, 6)
br$Speaker[grepl("Hobb", br$file) | grepl("Holi", br$file) | grepl("Home", br$file)] <- "Confederate"
br$Role[br$Speaker == "Confederate"] <- "Confederate"
br$Role[br$Speaker != "Confederate"] <- "Participant"

br[, c("Condition", "Speaker", "Role", "act")] <- lapply(br[, c("Condition", "Speaker", "Role", "act")], as.factor)
br$numberIPUs <- as.numeric(br$numberIPUs)

br$cycleDurZ <- (br$cycleDur - mean(br$cycleDur)) / sd(br$cycleDur)
br$breathCycleDurMeanZ <- (br$breathCycleDurMean - mean(br$breathCycleDurMean)) / sd(br$breathCycleDurMean)
br$breathRateZ <- (br$breathRate - mean(br$breathRate)) / sd(br$breathRate)
br$numberIPUsZ <- (br$numberIPUs - mean(br$numberIPUs, na.rm=TRUE)) / sd(br$numberIPUs, na.rm=TRUE)

br <- br %>%
  group_by(Role=="Participant") %>% # only delete the outliers of the participants; keep all of the confederate's data
  filter(abs(cycleDurZ) < 2) %>%
  filter(abs(breathCycleDurMeanZ) < 2) %>%
  filter(abs(breathRateZ) < 2) %>%
  filter(abs(numberIPUsZ) < 2 | is.na(numberIPUsZ))

orderCond <- c("Sitting", "Light", "Heavy")
orderCondBase <- c("Baseline", "Sitting", "Light", "Heavy")
orderAct <- c("watching", "listening", "speaking")

OGbr <- br
```

# Breathing rate (cycles per minute)

```{r, include=FALSE}
br <- br[!duplicated(br$file),]
```


## Per ACT (all conditions) 

```{r, echo=FALSE}
ggplot(br %>% filter(Role=="Participant"), aes(act, breathRate))+
  geom_boxplot()+
  ggtitle("breathing rate (across conditions)")+
  scale_x_discrete(limits = orderAct)

br$act <- relevel(br$act, ref="watching")
summary(lmer(breathRate ~ act + (1|Speaker), br %>% filter(Role == "Participant")))
```

## But dividing per condition, a similar pattern emerges every time, but with a difference between sitting and biking

### During SITTING

**No difference between watching and listening**
```{r, echo=FALSE}
ggplot(br %>% filter(Role=="Participant") %>% filter(Condition=="Sitting"), aes(act, breathRate))+
  geom_boxplot()+
  ggtitle("during SITTING")+
  scale_x_discrete(limits = orderAct)

summary(lmer(breathRate ~ act + (1|Speaker), br %>% filter(Role == "Participant") %>% filter(Condition=="Sitting")))
```

### During LIGHT BIKING

**Difference between watching and listening! Though not in the regression**
```{r, echo=FALSE}
ggplot(br %>% filter(Role=="Participant") %>% filter(Condition=="Light"), aes(act, breathRate))+
  geom_boxplot()+
  ggtitle("during LIGHT BIKING")+
  scale_x_discrete(limits = orderAct)

summary(lmer(breathRate ~ act + (1|Speaker), br %>% filter(Role == "Participant") %>% filter(Condition=="Light")))
```

### During HEAVY BIKING

**Also difference between watching and listening in graph but not in regression**
```{r, echo=FALSE}
ggplot(br %>% filter(Role=="Participant") %>% filter(Condition=="Heavy"), aes(act, breathRate))+
  geom_boxplot()+
  ggtitle("during HEAVY BIKING")+
  scale_x_discrete(limits = orderAct)

summary(lmer(breathRate ~ act + (1|Speaker), br %>% filter(Role == "Participant") %>% filter(Condition=="Heavy")))
```


## Per Condition (all acts)

```{r, echo=FALSE}
ggplot(br %>% filter(Role=="Participant"), aes(Condition, breathRate))+
  geom_boxplot()+
  scale_x_discrete(limits = orderCondBase)

br$Condition <- relevel(br$Condition, ref="Sitting")
summary(lmer(breathRate ~ Condition + (1|Speaker), br %>% filter(Role == "Participant")))
```

## Per condition during WATCHING 

### I guess we didn't replicate Paccalin & Jeannerod! (Though this could be a methodological problem: shorter recordings of "only watching")
```{r, echo=FALSE}
ggplot(br %>% filter(Role=="Participant") %>% filter(act=="watching"), aes(Condition, breathRate))+
  geom_boxplot()+
  ggtitle("during WATCHING (all in silence)")+
  scale_x_discrete(limits = orderCond)

summary(lmer(breathRate ~ Condition + (1|Speaker), br %>% filter(Role == "Participant") %>% filter(act=="watching")))
```

## per CONDITION during LISTENING

### there there seems to be a clear difference, but the regression says otherwise...

```{r, echo=FALSE}
ggplot(br %>% filter(Role=="Participant") %>% filter(act=="listening"), aes(Condition, breathRate))+
  geom_boxplot()+
  ggtitle("during LISTENING")+
  scale_x_discrete(limits = orderCond)

summary(lmer(breathRate ~ Condition + (1|Speaker), br %>% filter(Role == "Participant") %>% filter(act=="listening")))
```

## per CONDITION during SPEAKING

```{r, echo=FALSE}
ggplot(br %>% filter(Role=="Participant") %>% filter(act=="speaking"), aes(Condition, breathRate))+
  geom_boxplot()+
  ggtitle("during SPEAKING")+
  scale_x_discrete(limits = orderCondBase)

summary(lmer(breathRate ~ Condition + (1|Speaker), br %>% filter(Role == "Participant") %>% filter(act=="speaking")))
```
### but we can also exclude those 2 high values (z scores = 1.53 and 1.79), and then we find a slight effect between conditions, and a clearer effect in baseline vs conditions

```{r, echo=FALSE}
ggplot(br %>% filter(Role=="Participant") %>% filter(act=="speaking") %>% filter(breathRate < 17.5), aes(Condition, breathRate))+
  geom_boxplot()+
  ggtitle("during SPEAKING")+
  scale_x_discrete(limits = orderCondBase)

summary(lmer(breathRate ~ Condition + (1|Speaker), br %>% filter(Role == "Participant") %>% filter(act=="speaking") %>% filter(breathRate < 17.5)))
```

# 

## Interaction Condition * Act

**a bit hard to interpret**

```{r, echo=FALSE}
summary(lmer(breathRate ~ act * Condition + (1|Speaker), br %>% filter(Role == "Participant")))
```

# Breath cycle duration

```{r, include=FALSE}
br <- OGbr
```

## Confederate

**In confederate, expected change in breath cycle duration per condition (but in the regression there was no light vs heavy difference, although (the tendency to) the expeceted pattern still holds**
```{r, echo=FALSE}
ggplot(br %>% filter(Role=="Confederate"), aes(Condition, cycleDur))+
  geom_boxplot()+
  ggtitle("breath cycle duration (confederate)")+
  scale_x_discrete(limits = orderCond)

br$Condition <- relevel(br$Condition, ref="Sitting")
summary(lm(cycleDur ~ Condition, br %>% filter(Role == "Confederate")))

br$Condition <- relevel(br$Condition, ref="Light")
summary(lm(cycleDur ~ Condition, br %>% filter(Role == "Confederate")))
```

## Participants

## Per condition, during WATCHING

**no effect! Weird "outliers"?**


```{r, echo=FALSE}
ggplot(br %>% filter(Role=="Participant") %>% filter(act=="watching"), aes(Condition, cycleDur))+
  geom_boxplot()+
  ggtitle("breath cycle duration (during WATCHING)")+
  scale_x_discrete(limits = orderCond)

br$Condition <- relevel(br$Condition, ref="Sitting")
summary(lmer(cycleDur ~ Condition + (1|Speaker), br %>% filter(Role == "Participant") %>% filter(act=="watching")))
```

## Per condition, during LISTENING

**also no effect**
```{r, echo=FALSE}
ggplot(br %>% filter(Role=="Participant") %>% filter(act=="listening"), aes(Condition, cycleDur))+
  geom_boxplot()+
  ggtitle("breath cycle duration (during LISTENING)")+
  scale_x_discrete(limits = orderCond)

br$Condition <- relevel(br$Condition, ref="Sitting")
summary(lmer(cycleDur ~ Condition + (1|Speaker), br %>% filter(Role == "Participant") %>% filter(act=="listening")))
```

## Per condition, during SPEAKING

**also no effect**
```{r, echo=FALSE}
ggplot(br %>% filter(Role=="Participant") %>% filter(act=="speaking"), aes(Condition, cycleDur))+
  geom_boxplot()+
  ggtitle("breath cycle duration (during SPEAKING)")+
  scale_x_discrete(limits = orderCondBase)

br$Condition <- relevel(br$Condition, ref="Sitting")
summary(lmer(cycleDur ~ Condition + (1|Speaker), br %>% filter(Role == "Participant") %>% filter(act=="speaking")))
```


## Per ACT, across conditions

**shorter cycles during listening: makes sense with higher breath rate during listening (see above).**

**clear effects on graph and regression**
```{r, echo=FALSE}
ggplot(br %>% filter(Role=="Participant"), aes(act, cycleDur))+
  geom_boxplot()+
  ggtitle("breath cycle duration (across conditions)")+
  scale_x_discrete(limits = orderAct)

br$act <- relevel(br$act, ref="watching")
summary(lmer(cycleDur ~ act + (1|Speaker), br %>% filter(Role == "Participant")))
```

## Dividing by condition, cycle durations per act still hold the same pattern

## Per ACT, during SITTING

**watching vs listening: not so strong but present effect**
```{r, echo=FALSE}
ggplot(br %>% filter(Role=="Participant") %>% filter(Condition=="Sitting"), aes(act, cycleDur))+
  geom_boxplot()+
  ggtitle("breath cycle duration (during SITTING)")+
  scale_x_discrete(limits = orderAct)

br$act <- relevel(br$act, ref="watching")
summary(lmer(cycleDur ~ act + (1|Speaker), br %>% filter(Role == "Participant") %>% filter(Condition=="Sitting")))
```

## Per ACT, during LIGHT BIKING

**slightly stronger effect in watching vs listening**
```{r, echo=FALSE}
ggplot(br %>% filter(Role=="Participant") %>% filter(Condition=="Light"), aes(act, cycleDur))+
  geom_boxplot()+
  ggtitle("breath cycle duration (during LIGHT BIKING)")+
  scale_x_discrete(limits = orderAct)

br$act <- relevel(br$act, ref="watching")
summary(lmer(cycleDur ~ act + (1|Speaker), br %>% filter(Role == "Participant") %>% filter(Condition=="Light")))
```

## Per ACT, during HEAVY BIKING

**same watching vs listening effect as in light biking**
```{r, echo=FALSE}
ggplot(br %>% filter(Role=="Participant") %>% filter(Condition=="Heavy"), aes(act, cycleDur))+
  geom_boxplot()+
  ggtitle("breath cycle duration (during HEAVY BIKING)")+
  scale_x_discrete(limits = orderAct)

br$act <- relevel(br$act, ref="watching")
summary(lmer(cycleDur ~ act + (1|Speaker), br %>% filter(Role == "Participant") %>% filter(Condition=="Heavy")))
```


# Number of IPUs per breathing cycle, divded by condition

```{r, include=FALSE}
br <- OGbr
```

## Confederate

**When biking, the confederate seems to produce slightly fewer IPUs per breath cycle (makes sense! less breath when biking)**

```{r, echo=FALSE}
ggplot(br %>% filter(Role=="Confederate"), aes(Condition, numberIPUs))+
  geom_boxplot()+
  ggtitle("number of IPUs per breath cycle (confederate)")+
  scale_x_discrete(limits = orderCond)

br$Condition <- relevel(br$Condition, ref="Sitting")
summary(lm(numberIPUs ~ Condition, br %>% filter(Role == "Confederate")))

```
## Participants

**Participants really don't seem to change their IPUs per cycle across conditions**

```{r, echo=FALSE}
ggplot(br %>% filter(Role=="Participant"), aes(Condition, numberIPUs))+
  geom_boxplot()+
  ggtitle("number of IPUs per breath cycle (participants)")+
  scale_x_discrete(limits = orderCondBase)

summary(lmer(numberIPUs ~ Condition + (1|Speaker), br %>% filter(Role == "Participant")))
```

