---
title: "Confederate's Data"
output: html_document
---

```{r, include=FALSE}
library(tidyverse)
library(viridis)
library(cowplot)
library(lme4)

folder <- "C:/Users/tomof/Documents/1HU/ExperimentBreathing/Data/DataForAnalysis/AllData/"

order <- c("Sitting", "Light", "Heavy")
```

## The regressions below are simple and preliminary; they still have to be refined

# Free Speech

## Breathing rate

### Clear pattern for breathing rate during free speech

```{r, echo=FALSE}
load(paste0(folder, "DataNoDiff.RData"))
df <- dat %>% 
  filter(Speaker=="Confederate") %>%
  filter(Task=="Free")

ggplot()+
  geom_point(df, mapping=aes(x=Condition, y=breathRateSpeech), color="#2D708EFF", size=4)+
  ggtitle("Breathing rate (free speech)")+
  scale_x_discrete(limits = order)+
  theme(axis.title=element_text(size=20), axis.text=element_text(size=18), title=element_text(size=23))
```

**And this seems to be supported by a regression** (reference value: sitting)
```{r, echo=FALSE}
df$Condition <- relevel(df$Condition, ref="Sitting")
summary(lm(breathRateSpeech ~ Condition, data = df))
```

## f0

### f0 also changes clearly with condition

```{r, echo=FALSE}
load(paste0(folder, "DataNoDiff.RData"))
df <- dat %>% 
  filter(Speaker=="Confederate") %>%
  filter(Task=="Free")

ggplot()+
  geom_point(df, mapping=aes(x=Condition, y=f0mean), color="#2D708EFF", size=4)+
  ggtitle("f0 mean (free speech)")+
  scale_x_discrete(limits = order)+
  theme(axis.title=element_text(size=20), axis.text=element_text(size=18), title=element_text(size=23))
```



*Also supported by the regression** (reference value: sitting)
```{r, echo=FALSE}
df$Condition <- relevel(df$Condition, ref="Sitting")
summary(lm(f0mean ~ Condition, data = df))
```


### Same pattern if we divide by breathing rate instead of conditions

```{r, echo=FALSE}
ggplot() +
  geom_point(df, mapping=aes(x = breathRateSpeech, y = f0mean), color="#2D708EFF", size=4)+
  ggtitle("f0 (free speech)")+
  theme(legend.title=element_text(size=18), legend.text=element_text(size=16),
        axis.title=element_text(size=21), axis.text=element_text(size=20),
        plot.title=element_text(size=25))
```

**Also supported by regression**

```{r, echo=FALSE}
summary(lm(f0mean ~ breathRateSpeech, data = df))
```

## Speech rate

### Speech rate has this non-linear pattern

```{r, echo=FALSE}
load(paste0(folder, "DataNoDiff.RData"))
df <- dat %>% 
  filter(Speaker=="Confederate") %>%
  filter(Task=="Free")

ggplot()+
  geom_point(df, mapping=aes(x=Condition, y=speechRate), color="#2D708EFF", size=4)+
  ggtitle("Speech rate (free speech)")+
  scale_x_discrete(limits = order)+
  theme(axis.title=element_text(size=20), axis.text=element_text(size=18), title=element_text(size=23))
```

```{r, echo=FALSE}
##Artic Rate

load(paste0(folder, "DataNoDiff.RData"))
df <- dat %>% 
  filter(Speaker=="Confederate") %>%
  filter(Task=="Free")

ggplot()+
  geom_point(df, mapping=aes(x=Condition, y=articRate), color="#2D708EFF", size=4)+
  ggtitle("Articulation rate (free speech)")+
  scale_x_discrete(limits = order)+
  theme(axis.title=element_text(size=20), axis.text=element_text(size=18), title=element_text(size=23))
```

```{r, echo=FALSE}
##pause duration

load(paste0(folder, "DataNoDiff.RData"))
df <- dat %>% 
  filter(Speaker=="Confederate") %>%
  filter(Task=="Free")

df$pausePerc <-  (df$pauseDur*100)/df$durationSpeech

ggplot()+
  geom_point(df, mapping=aes(x=Condition, y=pausePerc), color="#2D708EFF", size=4)+
  ggtitle("PAuse duration perc (free speech)")+
  scale_x_discrete(limits = order)+
  theme(axis.title=element_text(size=20), axis.text=element_text(size=18), title=element_text(size=23))
```


**Regression suggests there's a clear difference between conditions** (reference value: sitting)
```{r, echo=FALSE}
df$Condition <- relevel(df$Condition, ref="Sitting")
summary(lm(f0mean ~ Condition, data = df))
```


### Same pattern if we divide by breathing rate instead of conditions

```{r, echo=FALSE}
ggplot() +
  geom_point(df, mapping=aes(x = breathRateSpeech, y = speechRate), color="#2D708EFF", size=4)+
  ggtitle("Speech rate (free speech)")+
  theme(legend.title=element_text(size=18), legend.text=element_text(size=16),
        axis.title=element_text(size=21), axis.text=element_text(size=20),
        plot.title=element_text(size=25))
```

**Also supported by regression**

```{r, echo=FALSE}
summary(lm(f0mean ~ breathRateSpeech, data = df))
```

# Read Speech

## Breathing rate

### During read speech breathing rate has this weird pattern.

```{r, echo=FALSE}
load(paste0(folder, "DataNoDiff.RData"))
df <- dat %>% 
  filter(Speaker=="Confederate") %>%
  filter(Task=="ReadAlone")

ggplot()+
  geom_point(df, mapping=aes(x=Condition, y=breathRateSpeech), color="#2D708EFF", size=4)+
  ggtitle("Breathing rate (read speech)")+
  scale_x_discrete(limits = order)+
  theme(axis.title=element_text(size=20), axis.text=element_text(size=18), title=element_text(size=23))
```

### Divided by text

```{r, echo=FALSE}

ggplot()+
  geom_point(df, mapping=aes(x=Condition, y=breathRateSpeech), color="#2D708EFF", size=4)+
  ggtitle("Breathing rate (read speech)")+
  scale_x_discrete(limits = order)+
  facet_wrap(~Topic)+
  theme(axis.title=element_text(size=20), axis.text=element_text(size=18), title=element_text(size=23))
```


**Regression shows nothing** (reference value: sitting)
```{r, echo=FALSE}
df$Condition <- relevel(df$Condition, ref="Sitting")
summary(lm(breathRateSpeech ~ Condition, data = df))
```

## f0

### But f0 changes clearly with condition, so the physical effort did influence f0

```{r, echo=FALSE}
load(paste0(folder, "DataNoDiff.RData"))
df <- dat %>% 
  filter(Speaker=="Confederate") %>%
  filter(Task=="ReadAlone")

ggplot()+
  geom_point(df, mapping=aes(x=Condition, y=f0mean), color="#2D708EFF", size=4)+
  ggtitle("f0 mean (read speech)")+
  scale_x_discrete(limits = order)+
  theme(axis.title=element_text(size=20), axis.text=element_text(size=18), title=element_text(size=23))
```



*Also supported by the regression** (reference value: sitting)
```{r, echo=FALSE}
df$Condition <- relevel(df$Condition, ref="Sitting")
summary(lm(f0mean ~ Condition, data = df))
```


### But since here breathing rate != condition, f0 doesn't follow breathing rate

```{r, echo=FALSE}
ggplot() +
  geom_point(df, mapping=aes(x = breathRateSpeech, y = f0mean), color="#2D708EFF", size=4)+
  ggtitle("f0 (read speech)")+
  theme(legend.title=element_text(size=18), legend.text=element_text(size=16),
        axis.title=element_text(size=21), axis.text=element_text(size=20),
        plot.title=element_text(size=25))
```

**Also in the regression**

```{r, echo=FALSE}
summary(lm(f0mean ~ breathRateSpeech, data = df))
```

## Speech rate

### Speech rate goes down with heavier conditions

```{r, echo=FALSE}
load(paste0(folder, "DataNoDiff.RData"))
df <- dat %>% 
  filter(Speaker=="Confederate") %>%
  filter(Task=="ReadAlone")

ggplot()+
  geom_point(df, mapping=aes(x=Condition, y=speechRate), color="#2D708EFF", size=4)+
  ggtitle("Speech rate (read speech)")+
  scale_x_discrete(limits = order)+
  theme(axis.title=element_text(size=20), axis.text=element_text(size=18), title=element_text(size=23))
```



**Regression suggests it's significant** (reference value: sitting)
```{r, echo=FALSE}
df$Condition <- relevel(df$Condition, ref="Sitting")
summary(lm(f0mean ~ Condition, data = df))
```


### Again, dividing by breathing rate doesn't show anything

```{r, echo=FALSE}
ggplot() +
  geom_point(df, mapping=aes(x = breathRateSpeech, y = speechRate), color="#2D708EFF", size=4)+
  ggtitle("Speech rate (read speech)")+
  theme(legend.title=element_text(size=18), legend.text=element_text(size=16),
        axis.title=element_text(size=21), axis.text=element_text(size=20),
        plot.title=element_text(size=25))
```

**Also in the regression**

```{r, echo=FALSE}
summary(lm(f0mean ~ breathRateSpeech, data = df))
```