---
title: "Breathing Rate in Joint Reading"
output: html_document
---

```{r, include=FALSE}
library(tidyverse)
library(viridis)
library(cowplot)
library(lme4)

folder <- "C:/Users/tomof/Documents/1HU/ExperimentBreathing/Data/DataForAnalysis/AllData/"

order <- c("Baseline", "Sitting", "Light", "Heavy")
orderconf <- order[-1]
```

## The regressions below are simple and preliminary; they still have to be refined

### Looking at conditions: maybe slight trend upwards with condition

```{r, echo=FALSE}
load(paste0(folder, "DataNoDiff.RData"))
df <- dat[dat$Task=="ReadJoint",]

ggplot()+
  geom_boxplot(df %>% filter(Role=="Participant"), mapping=aes(Condition, breathRateSpeech), fill="#2D708EFF")+ # I got this color code from a viridis palette with scales::show_col(viridis_pal(option="D)(12))
  geom_point(df %>% filter(Role=="Confederate"), mapping=aes(x=Condition, y=breathRateSpeech), color="#7AD151FF", size=5)+
  scale_colour_manual(values=c("Participants"="#2D708EFF", "Confederate"="#7AD151FF"))+
  ggtitle("breathing rate (ReadJoint)")+
  scale_x_discrete(limits = orderconf)+
  theme(axis.title=element_text(size=20), axis.text=element_text(size=18), title=element_text(size=23))

```



**In the regression: small trend upwards (no difference between sitting and light or between light and heavy, but difference between sitting and heavy.** (reference value: sitting)
```{r, echo=FALSE}
df$Condition <- relevel(df$Condition, ref="Light")
summary(lmer(breathRateSpeech ~ Condition + (1 | Speaker), data = df))
```


### Seeing breathing rate along confederate's breathing rate, I don't see much difference

```{r, echo=FALSE}
load(paste0(folder, "DataNoDiff.RData"))

conf <- dat[dat$Speaker=="Confederate", c(3, 4, 5, 7:13, 26, 28:31, 33)]
colnames(conf) <- sub("^","C", colnames(conf))
dat2 <-dat
dat3 <- dat2[rep(seq_len(nrow(dat2)), each = nrow(conf)),]
rownames(dat3) <- 1:nrow(dat3)
conf2 <- do.call("rbind", replicate((nrow(dat3)/nrow(conf)), conf, simplify=FALSE))

dat4 <- cbind(dat3, conf2)

dat4 <- dat4[(dat4$Condition == dat4$CCondition & dat4$Task == dat4$CTask) | (dat4$Condition == "Baseline"),]
dat4[dat4$Condition == "Baseline", colnames(conf)] <- NA
dat4 <- dat4[!duplicated(dat4$file),]

df <- dat4 %>% filter(Task=="ReadJoint")

ggplot() +
  geom_point(df %>% filter(Role=="Participant"),
    mapping=aes(x = CbreathRateSpeech, y = breathRateSpeech, color = Role), size=4) +
  geom_point(df %>% filter(Role=="Confederate"),
             mapping=aes(x = CbreathRateSpeech, y = breathRateSpeech, color = Role), size=4)+
  scale_color_viridis_d("Speaker", direction=-1, end=.65)+
  ggtitle("breathing rate (ReadJoint)")+
  theme(legend.title=element_text(size=18), legend.text=element_text(size=16),
        axis.title=element_text(size=21), axis.text=element_text(size=20),
        plot.title=element_text(size=25))
```

**The regression also doesn't support a difference following the confederate's breathing rate** (reference value: baseline)
```{r, echo=FALSE}
load(paste0(folder, "DataWithDiff.RData"))
df <- dat[dat$Task == "ReadJoint",]
summary(lmer(breathRateSpeech ~ CbreathRateSpeech + (1 | Speaker), data = df))
```


## We can also look at the difference of participants' minus confederate's f0:

### Looking at categorical conditions: not sure how to interpret these differences. Biggest difference in heavy condition? (makes sense: confederate changed a lot, but participants don't follow as much)

```{r, echo=FALSE}
load(paste0(folder, "DataWithDiff.RData"))
df <- dat[dat$Task=="ReadJoint" & dat$Condition!="Baseline",]

ggplot(df, aes(Condition, breathRateSpeechDiff))+
  geom_boxplot(fill="#2D708EFF")+ # I got this color code from a viridis palette with scales::show_col(viridis_pal(option="D)(12))
  ggtitle("breath rate difference (participant - confederate) (ReadJoint)")+
  scale_x_discrete(limits = orderconf)+
  theme(axis.title=element_text(size=20), axis.text=element_text(size=18),
        title=element_text(size=23))
```

**This difference between conditions seems clear in the regression, but because confederate changed a lot.** (Reference value: sitting)

```{r, echo=FALSE}
load(paste0(folder, "DataWithDiff.RData"))
df <- dat[dat$Task == "ReadJoint",]
df$Condition <- relevel(df$Condition, ref="Sitting")
summary(lmer(breathRateSpeechDiff ~ Condition + (1 | Speaker), data = df))
```

### Looking at difference according to confederate's breathing rate, it's the same pattern as above

```{r, echo=FALSE}
load(paste0(folder, "DataWithDiff.RData"))

df <- dat %>% filter(Task=="ReadJoint")

ggplot(df, aes(CbreathRateSpeech, breathRateSpeechDiff))+
  geom_point(color="#2D708EFF")+ # I got this color code from a viridis palette with scales::show_col(viridis_pal(option="D)(12))
  ggtitle("breath rate difference (participant - confederate) (ReadJoint)")+
  theme(axis.title=element_text(size=20), axis.text=element_text(size=18),
        title=element_text(size=23))
```

**Regression shows slight decrease in difference with breathing rate.**

```{r, echo=FALSE}
load(paste0(folder, "DataWithDiff.RData"))
df <- dat %>% filter(Task=="ReadJoint")
summary(lmer(breathRateSpeechDiff ~ CbreathRateSpeech + (1 | Speaker), data = df))

```


# Conclusion: we need more data?
