---
title: "Speech Rate in Joint Reading"
output: html_document
---

```{r, include=FALSE}
library(tidyverse)
library(viridis)
library(cowplot)
library(lme4)

folder <- "C:/Users/tomof/Documents/1HU/ExperimentBreathing/Data/DataForAnalysis/AllData/"


order <- c("Baseline", "Sitting", "Light", "Heavy")
orderconf <- order[-1]
```

## The regressions below are simple and preliminary; they still have to be refined

### Looking at conditions, people seem to have followed the confederate's speech rate

```{r, echo=FALSE}
load(paste0(folder, "DataNoDiff.RData"))
df <- dat[dat$Task=="ReadJoint",]

ggplot()+
  geom_boxplot(df %>% filter(Role=="Participant"), mapping=aes(Condition, speechRate), fill="#2D708EFF")+ # I got this color code from a viridis palette with scales::show_col(viridis_pal(option="D)(12))
  geom_point(df %>% filter(Role=="Confederate"), mapping=aes(x=Condition, y=speechRate), color="#7AD151FF", size=5)+
  scale_colour_manual(values=c("Participants"="#2D708EFF", "Confederate"="#7AD151FF"))+
  ggtitle("speech rate (ReadJoint speech)")+
  scale_x_discrete(limits = orderconf)+
  theme(axis.title=element_text(size=20), axis.text=element_text(size=18), title=element_text(size=23))

```
### Divided per text

```{r, echo=FALSE}
ggplot()+
  geom_boxplot(df %>% filter(Role=="Participant"), mapping=aes(Condition, speechRate), fill="#2D708EFF")+ # I got this color code from a viridis palette with scales::show_col(viridis_pal(option="D)(12))
  geom_point(df %>% filter(Role=="Confederate"), mapping=aes(x=Condition, y=speechRate), color="#7AD151FF", size=5)+
  scale_colour_manual(values=c("Participants"="#2D708EFF", "Confederate"="#7AD151FF"))+
  ggtitle("speech rate (ReadJoint speech)")+
  scale_x_discrete(limits = orderconf)+
  facet_wrap(~Topic)+
  theme(axis.title=element_text(size=20), axis.text=element_text(size=18), title=element_text(size=23))

```

**The regression supports this difference** (reference value: sitting)
```{r, echo=FALSE}
df$Condition <- relevel(df$Condition, ref="Sitting")
summary(lmer(speechRate ~ Condition + (1 | Speaker), data = df))
```


### If we look at confederate's speech rate instead of categorical conditions: same weirdness in the non-linearity, but the participants still seem to follow the confederate's speech rate

```{r, echo=FALSE}
load(paste0(folder, "DataNoDiff.RData"))

conf <- dat[dat$Speaker=="Confederate", c(3, 4, 5, 7:13, 26, 28:31, 33)]
colnames(conf) <- sub("^","C", colnames(conf))
dat2 <-dat
dat3 <- dat2[rep(seq_len(nrow(dat2)), each = nrow(conf)),]
rownames(dat3) <- 1:nrow(dat3)
conf2 <- do.call("rbind", replicate((nrow(dat3)/nrow(conf)), conf, simplify=FALSE))

dat4 <- cbind(dat3, conf2)

dat4 <- dat4[(dat4$Condition == dat4$CCondition & dat4$Task == dat4$CTask) | (dat4$Condition == "Baseline"),]
dat4[dat4$Condition=="Baseline", colnames(conf)] <- NA
dat4 <- dat4[!duplicated(dat4$file),]

df <- dat4 %>%
  filter(Task=="ReadJoint")

ggplot() +
  geom_point(df %>% filter(Role=="Participant"),
    mapping=aes(x = CbreathRateSpeech, y = speechRate, color = Role), size=4) +
  geom_point(df %>% filter(Role=="Confederate"),
             mapping=aes(x = CbreathRateSpeech, y = speechRate, color = Role), size=4)+
  scale_color_viridis_d("Speaker", direction=-1, end=.65)+
  ggtitle("speech rate (ReadJoint speech)")+
  theme(legend.title=element_text(size=18), legend.text=element_text(size=16),
        axis.title=element_text(size=21), axis.text=element_text(size=20),
        plot.title=element_text(size=25))
```

**The regression seems to support it**
```{r, echo=FALSE}
load(paste0(folder, "DataWithDiff.RData"))
df <- dat[dat$Task == "ReadJoint",]
summary(lmer(speechRate ~ CbreathRateSpeech + (1 | Speaker), data = df))
```


## We can also look at the difference of participants' minus confederate's f0:

### Looking at categorical conditions: weird non-linearity. Light biking: confederate spoke faster; sitting and heavy: participants spoke faster.

```{r, echo=FALSE}
load(paste0(folder, "DataWithDiff.RData"))
df <- dat[dat$Task=="ReadJoint" & dat$Condition!="Baseline",]

ggplot(df, aes(Condition, speechRateDiff))+
  geom_boxplot(fill="#2D708EFF")+ # I got this color code from a viridis palette with scales::show_col(viridis_pal(option="D)(12))
  ggtitle("speech rate difference (participant - confederate) (ReadJoint speech)")+
  scale_x_discrete(limits = orderconf)+
  theme(axis.title=element_text(size=20), axis.text=element_text(size=18),
        title=element_text(size=23))
```

**The regression confirms this difference between conditions.** (Reference value: sitting)

```{r, echo=FALSE}
load(paste0(folder, "DataWithDiff.RData"))
df <- dat[dat$Task == "ReadJoint",]
df$Condition <- relevel(df$Condition, ref="Sitting")
summary(lmer(speechRateDiff ~ Condition + (1 | Speaker), data = df))
```

## Looking at difference according to confederate's breathing rate, again weird trend

```{r, echo=FALSE}
load(paste0(folder, "DataWithDiff.RData"))

df <- dat %>%
  filter(Task=="ReadJoint")

ggplot(df, aes(CbreathRateSpeech, speechRateDiff))+
  geom_point(color="#2D708EFF")+ # I got this color code from a viridis palette with scales::show_col(viridis_pal(option="D)(12))
  ggtitle("speech rate difference (participant - confederate) (ReadJoint speech)")+
  theme(axis.title=element_text(size=20), axis.text=element_text(size=18),
        title=element_text(size=23))
```


**The regression seems to suggest a slight trend, but again, it's non-linear so it doesn't capture it well.**

```{r, echo=FALSE}
load(paste0(folder, "DataWithDiff.RData"))
df <- dat %>% filter(Task=="ReadJoint")
summary(lmer(speechRateDiff ~ CbreathRateSpeech + (1 | Speaker), data = df))

```


# Conclusion: we need more data?
